\documentclass{amsart}
\usepackage{amsmath,amssymb,latexsym,enumerate}
\usepackage{pstricks}
\usepackage{pst-node}
\usepackage{tikz}


\newtheorem{theorem}{Theorem}[section]
\newtheorem{subtheorem}{Theorem}[theorem]
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{subproposition}{Proposition}[theorem]
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{subexample}{Example}[theorem]
\newtheorem{subexercise}{Exercise}[theorem]
\newtheorem{exercise}[theorem]{Exercise}
\theoremstyle{remark} 
\newtheorem{remark}[theorem]{Remark}
\numberwithin{equation}{section}


\newcommand{\cA}{{\mathcal{A}}}
\newcommand{\cF}{{\mathcal{F}}}
\newcommand{\cL}{{\mathcal{L}}}
\newcommand{\cO}{{\mathcal{O}}}
\newcommand{\cT}{{\mathcal{T}}}
\newcommand{\cU}{{\mathcal{U}}}
\newcommand{\ba}{{\bf a}}
\newcommand{\bb}{{\bf b}}
\newcommand{\bff}{{\bf f}}
\newcommand{\x}{{\bf x}}
\newcommand{\bx}{{\bf x}}
\newcommand{\y}{{\bf y}}
\newcommand{\by}{{\bf y}}
\newcommand{\CC}{{\mathbb{C}}}
\newcommand{\PP}{{\mathbb{P}}}
\newcommand{\ZZ}{{\mathbb{Z}}}

\newcommand{\meet}[2]{#1 \cap #2}
%\newcommand{\join}[2]{\overleftrightarrow{#1 #2}}
\newcommand{\join}[2]{\langle #1, #2 \rangle}
\newcommand{\sgn}{\operatorname{sign}}

\newcommand{\erase}[1]{{}}


\title{Introduction to Cluster Algebras}
\author{Max Glick}
\author{Dylan Rupel}

\begin{document}
\begin{abstract}
  These are notes for a series of lectures presented at the ASIDE conference 2016.
\end{abstract}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
  Cluster algebras were introduced by Fomin and Zelevinsky \cite{FZ02} in 2002 as the culmination of their study of total positivity \cite{FZ99} and (dual) canonical bases.  The topic of cluster algebras quickly grew into its own as a subject deserving independent study mainly fueled by its emergent close relationship to many areas of mathematics.  Here is a partial list of related topics: combinatorics \cite{MP07}, hyperbolic geometry \cite{FG06,FST08,MSW13}, Lie theory \cite{GLS06}, Poisson geometry \cite{GSV03}, integrable systems \cite{dFK10,G11}, representations of associative algebras \cite{CC06,CK06,BMRRT06,R11,Q12,R15}, mathematical physics \cite{EF12,ABCGPT14}, and quantum groups \cite{K12,GLS13,KQ14,BR15}.

  In these notes we will give an introduction to cluster algebras and a couple of the applications mentioned above.  These notes are far from exhaustive and the references above only touch on the vast literature.  Other overviews of cluster algebras can be found in the works \cite{K10,W13,GLS} which will also provide additional references.

  These lecture notes are organized as follows.  Section~\ref{sec:motivation} gives several motivating examples from which we will abstract the definition of a cluster algebra.  %The reader will be guided towards providing ad-hoc proofs of some of the important theorems of cluster algebra theory in the exercises of that section.  
	Section~\ref{sec:cluster_algebras} contains several variations on the definition of cluster algebras with increasing generality and also the definition of $Y$-patterns.  In Section~\ref{sec:basic_theorems}, we describe the foundational results in the theory of cluster algebras and sketch, or otherwise indicate the ideas behind, the proofs of these results. Section~\ref{sec:poisson_and_quantum} recalls the theory of Poisson structures compatible with a cluster algebra and describes how this naturally leads to a quantization of cluster algebras.  Finally, we conclude with applications of the cluster algebra machinery to problems involving integrable systems in Section~\ref{sec:integrable_systems}.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Motivating Examples}\label{sec:motivation}
	Cluster algebras are certain commutative rings possessing additional structure, including:
	\begin{itemize}
		\item a distinguished collection of generators called \emph{cluster variables};
		\item a collection of finite subsets of the set of all cluster variables called \emph{clusters};
		\item a \emph{mutation} rule which, given a cluster and one of its variables $x$, produces another cluster by replacing $x$ with a different cluster variable $x'$ related to $x$ by an \emph{exchange relation}
		\begin{displaymath}
			xx' = F,
		\end{displaymath}
		where $F$ is a binomial in the variables common to the initial and mutated clusters.
	\end{itemize}
	Before defining cluster algebras, we describe a few settings illustrating these various components.  We leave it to the astute reader to compare these examples with the definitions presented in Section~\ref{sec:cluster_algebras} and to see how they may be reinterpreted within the cluster algebra framework.

	\subsection{Markov Triples} \label{sub:Markov triples}
  %\begin{example}\label{example:Markov triples}
    A Markov triple is a tuple $(a,b,c)$ of positive integers satisfying the Markov equation $a^2+b^2+c^2=3abc$; an integer which appears as a term in a Markov triple is called a Markov number.  The Markov equation is an example of a Diophantine equation and two classical number theoretic problems are to determine the number of solutions and to determine a method for finding all such solutions.  We will solve both of these problems for the Markov equation.
    
    \begin{exercise}
      Prove that $(1,1,1)$ and $(1,1,2)$ are (up to reordering) the only Markov triples with repeated values.
    \end{exercise}
    Rearranging the Markov equation we see that $c^2-3abc+a^2+b^2=0$ and so $c$ is a root of the quadratic $f(x)=x^2-(3ab)x+(a^2+b^2)$.  But notice that the other root $c'=3ab-c=\frac{a^2+b^2}{c}$ is a positive integer and so $(a,b,c')$ is again a Markov triple.  

    Note that there was nothing special about $c$ in the calculation above so that, given any Markov triple $(a,b,c)$ we may perform three possible exchanges 
    \[\Big(a,b,\frac{a^2+b^2}{c}\Big)\quad\Big(a,\frac{a^2+c^2}{b},c\Big)\quad\Big(\frac{b^2+c^2}{a},b,c\Big)\]
    and obtain another Markov triple in each case.  The following exercises solve the above two classical problems of Diophantine equations.
    \begin{exercise}\mbox{}
      \begin{enumerate}[\quad\upshape (a)]
        \item Prove that there are infinitely many Markov triples by showing that there is no bound on how large the largest value can be.
        \item Show that all Markov triples may be obtained from the Markov triple $(1,1,1)$ by a sequence of exchanges.
      \end{enumerate}
    \end{exercise}
  %\end{example}

	\subsection{The Grassmannian $Gr_{2}(\CC^n)$} \label{sub:Gr2n}
  %\begin{example} \label{example:Gr2n}
		The \emph{Grassmannian} $Gr_{k}(\CC^n)$ is the set of $k$-dimensional linear subspaces of $\CC^n$.  A point in the Grassmannian can be described, albeit non-uniquely, as the row span of a full rank matrix $A \in \CC^{k \times n}$.  The maximal minors of $A$ are called \emph{Pl\"ucker coordinates}.
		
		Now, restrict to $k=2$ and let 
		\begin{displaymath}
			A = \left[ \begin{array}{cccc}
			a_{11} & a_{12} & \ldots & a_{1n} \\
			a_{21} & a_{22} & \ldots & a_{2n} \\
			\end{array} \right].
		\end{displaymath}
		The Pl\"ucker coordinates are given by $\Delta_{ij} = a_{1i}a_{2j} - a_{1j}a_{2i}$ for $1 \leq i < j \leq n$.  As the matrix $A$ is not uniquely determined by a point in $Gr_2(\CC^n)$, the Pl\"ucker coordinates are not truly functions on $Gr_2(\CC^n)$, but rather they are only well-defined functions up to simultaneous rescaling.  That is, the Pl\"ucker coordinates determine a closed embedding into the projective space $\PP^{{n\choose 2}-1}$.  A first step in establishing this claim is given by the following exercise.
		\begin{exercise}
			Verify that the $\Delta_{ij}$ satisfy the so-called \emph{Pl\"ucker relations}
			\begin{equation}\label{eq:short plucker special}
				\Delta_{ik}\Delta_{jl} = \Delta_{ij}\Delta_{kl} + \Delta_{il}\Delta_{jk}
			\end{equation}
			for $1 \leq i < j < k < l \leq n$.
		\end{exercise}
		
		Consider a regular $n$-gon with vertices labelled $1,2,\ldots, n$.  Let $T$ be a triangulation, i.e. a maximal collection of chords $\overline{ij}$ with $1 \leq i < j \leq n$, no two of which intersect in their interiors.  Note that $T$ always consists of $n-3$ diagonals together with the $n$ sides $\overline{12}, \overline{23}, \overline{34}, \ldots, \overline{1n}$.  Associate to $T$ the corresponding collection of Pl\"ucker coordinates
		\begin{displaymath}
			\Delta_T := \{\Delta_{ij} : \overline{ij} \in T\}.
		\end{displaymath}
    \setcounter{subproposition}{1}
		\begin{proposition} \label{prop:Gr2n}
		Fix positive reals $x_{ij}$ for all $\overline{ij} \in T$.  Then there exists $A \in Gr_2(\CC^n)$ such that $\Delta_{ij}(A) = x_{ij}$ for all $\overline{ij} \in T$.  Moreover, each $\Delta_{ij}(A)$ with $\overline{ij} \notin T$ can be expressed as a subtraction free rational expression of the $x_{ij}$.
		\end{proposition}
		
		For instance, let $n=4$ and $T = \{\overline{13},\overline{12},\overline{23},\overline{34},\overline{14}\}$.  A possible representing matrix is
		\begin{displaymath}
			A = \left[\begin{array}{cccc}
			1 & \frac{x_{23}}{x_{13}} & 0 & -\frac{x_{34}}{x_{13}} \\
			0 & x_{12} & x_{13} & x_{14} \\
			\end{array}\right].
		\end{displaymath}
		There is one remaining Pl\"ucker coordinate, namely $\Delta_{24}$, and it can be computed from the given ones as
		\begin{displaymath}
		\Delta_{24}(A) = \frac{x_{12}x_{34} + x_{14}x_{23}}{x_{13}}
		\end{displaymath}
		which follows from a Pl\"ucker relation.	This formula can be thought of as a change of coordinates from $\Delta_T$ to $\Delta_{T'}$ where $T' = (T \setminus \{\overline{13}\}) \cup \{\overline{24}\}$.  
    \setcounter{subexercise}{2}
    \begin{exercise}
      The change of coordinate systems observed above can be understood more classically in terms of the Ptolemy relations for cyclic quadrilaterals.  Indeed, consider four distinct points labelled $1,2,3,4$ on a circle in cyclic order and let $x_{ij}$ denote the distance between vertices $i$ and $j$.  Prove that $x_{13}x_{24}=x_{12}x_{34}+x_{14}x_{23}$.
    \end{exercise}
		
		More generally, if $T$ is a triangulation of an $n$-gon and $\overline{ik} \in T$ is a diagonal (as opposed to a side), then $\overline{ik}$ is part of two triangles of $T$.  Call the third vertices of these two triangles $j$ and $l$.  It follows that $T' = (T \setminus \{\overline{ik}\}) \cup \{\overline{jl}\}$ is a triangulation.  Performing a sequence of such quadrilateral flips and using a Pl\"ucker relation at each step makes it possible to iteratively compute all the rational expressions promised in Proposition \ref{prop:Gr2n}.
%  \end{example}
 
	\subsection{Double Bruhat cells} \label{sub:total positivity}
%  \begin{example}\label{example:total positivity}
    An $n\times n$ matrix $M$ is called \emph{totally positive} if the determinant of every square submatrix is a positive real number.  In particular, every entry of $M$ is positive and $M$ must be invertible.  Write $GL_n^{>0}\subset GL_n$ for the subset of totally positive matrices.  To check that a given matrix $M\in GL_n$ is totally positive one must, a priori, check that all ${2n\choose n}-1$ minors of $M$ are positive.  A natural question is whether this verification process can be made more efficient.  More precisely, is there a smaller collection of minors one may compute and from the positivity of this subset conclude that every minor is positive, i.e. conclude that $M\in GL_n^{>0}$?  We will call such a collection a \emph{total positivity criterion} if it exists.  

    For small $n$, such criteria can be easily found and easily verified.  For example, a matrix $M=\left[\begin{array}{cc} a & b\\ c & d\end{array}\right]\in GL_2$ is totally positive if and only if $a,b,c,ad-bc>0$ if and only if $b,c,d,ad-bc>0$.
    \begin{exercise}
      Find a minimal collection of minors whose positivity guarantees a matrix in $GL_3$ is totally positive (hint: any such total positivity criterion consists of 9 minors).
    \end{exercise}

    To describe a solution and easily identify total positivity criteria for all general linear groups $GL_n$, it will be convenient to slightly generalize the notion of total positivity.  An $n\times n$ matrix $M$ is called \emph{totally nonnegative} if the determinant of every square submatrix is a nonnegative real number.  Write $GL_n^{\ge0}\subset GL_n$ for the subset of totally nonnegative matrices.  Again one may ask: what is a minimal collection of minors needed to check that a matrix is totally nonnegative?  Unfortunately, or perhaps fortunately, the total nonnegativity criteria are not uniformly described across all of $GL_n$.  The solution to this problem naturally leads one to study certain subvarieties of $GL_n$ called \emph{double Bruhat cells}, which we now describe.

    Let $B_+,B_-\subset GL_n$ denote the subgroups of upper and lower triangular matrices respectively.  Identify the symmetric group $\Sigma_n$ with the subgroup of $GL_n$ consisting of permutation matrices, i.e. matrices having precisely one nonzero entry 1 in each row and column.  For example, identify the permutation $(1,2)\in\Sigma_2$ (written in cycle notation) with the matrix $\left[\begin{array}{cc}0 & 1\\ 1 & 0\end{array}\right]$.  

    It is well known that $GL_n$ decomposes in two ways (actually many ways) as a union of \emph{Bruhat cells}:
    \[GL_n=\bigsqcup_{w\in\Sigma_n}B_+w B_+=\bigsqcup_{w\in\Sigma_n}B_-w B_-.\]
    To understand total nonnegativity criteria for $GL_n$, we will consider the \emph{double Bruhat cells} $GL_n^{u,v}=B_+uB_+\cap B_-vB_-$ for $u,v\in\Sigma_n$.  It turns out that each double Bruhat cell admits its own collection of total positivity criteria, i.e. for each $u,v\in\Sigma_n$ there exists a minimal collection of minors whose positivity identifies the subset $GL_n^{u,v}\cap GL_n^{\ge0}$ inside the double Bruhat cell $GL_n^{u,v}$.

    For $I,J\subset[1,n]$ with $|I|=|J|$, denote by $\Delta_{I,J}$ the function on $GL_n$ which returns the determinant of the submatrix on row set $I$ and column set $J$.
    \setcounter{subexample}{1}
    \begin{example}\label{example:big cell}
      Let $w_0=(1,n)(2,n-1)\cdots(\lfloor n/2\rfloor,\lceil n/2\rceil)$ denote the longest permutation in $\Sigma_n$.  The double Bruhat cell $GL_n^{w_0,w_0}$ is given as follows:
      \[GL_n^{w_0,w_0}=\{M\in GL_n:\Delta_{[1,i],[n+1-i,n]}(M)\ne0,\Delta_{[n+1-i,n],[1,i]}(M)\ne0\text{ for all $i$}\}.\]
    \end{example}
    \setcounter{subexercise}{2}
    \begin{exercise}
      Find all 4 double Bruhat cells in $GL_2$ and verify that they partition the space.
    \end{exercise}

    Some of these total nonnegativity criteria in $GL_n^{u,v}$ can be conveniently described using double wiring diagrams.  Write $s_i=(i,i+1)\in\Sigma_n$ for the simple transposition interchanging $i$ and $i+1$.  A \emph{reduced word} for $w\in\Sigma_n$ is a minimal sequence $(i_1,\ldots,i_r)$ so that $w=s_{i_1}\cdots s_{i_r}$, where $\ell(w):=r$ is called the \emph{length} of $w$.  A reduced word $(i_1,\ldots,i_r)$ is naturally encoded in a \emph{wiring diagram}, i.e. a collection of $n$ strands with a crossing between the $i$-th and $(i+1)$-st strands each time $i$ appears in the reduced word, see Figure~\ref{fig:wiring diagram}.
    \begin{figure}[ht]
    \begin{tikzpicture}
      \draw (0,0) -- (0.5,0) -- (1.5,1) -- (2,1) -- (3,2) -- (3.5,2) -- (4.5,3) -- (5,3) -- (6,3) -- (6.5,3) -- (7.5,3) -- (8,3) -- (9,3) -- (9.5,3);
      \draw (0,1) -- (0.5,1) -- (1.5,0) -- (2,0) -- (3,0) -- (3.5,0) -- (4.5,0) -- (5,0) -- (6,1) -- (6.5,1) -- (7.5,2) -- (8,2) -- (9,2) -- (9.5,2);
      \draw (0,2) -- (0.5,2) -- (1.5,2) -- (2,2) -- (3,1) -- (3.5,1) -- (4.5,1) -- (5,1) -- (6,0) -- (6.5,0) -- (7.5,0) -- (8,0) -- (9,1) -- (9.5,1);
      \draw (0,3) -- (0.5,3) -- (1.5,3) -- (2,3) -- (3,3) -- (3.5,3) -- (4.5,2) -- (5,2) -- (6,2) -- (6.5,2) -- (7.5,1) -- (8,1) -- (9,0) -- (9.5,0);
    \end{tikzpicture}
    \caption{A wiring diagram for the reduced word $(1,2,3,1,2,1)\in\Sigma_4$.}
    \label{fig:wiring diagram}
    \end{figure}

    Given $(u,v)\in\Sigma_n^2$, a double reduced for $(u,v)$ is an arbitrary shuffle of a reduced word for $u$ and a reduced word for $v$, where terms in the reduced word for $v$ are taken from the set $\{-1,\ldots,-(n-1)\}$ for clarity.  For example, taking $u=(132),v=(13)\in\Sigma_3$, one double reduced word for $(u,v)$ is $(-1,2,-2,1,-1)$.  Given a double reduced word $(i_1,\ldots,i_r)$, build a double wiring diagram by superposing wiring diagrams associated to the reduced word for $u$ and the reduced word for $v$ with crossings ordered according to the double reduced word, see Figure~\ref{fig:double wiring diagram}.  
    \begin{figure}[ht]
    \begin{tikzpicture}
      %u
      \draw (0,0) -- (2,0) -- (3,0) -- (5,0) -- (6,1) -- (8,1) -- (9,2) -- (11,2);
      \draw (0,1) -- (2,1) -- (3,2) -- (5,2) -- (6,2) -- (8,2) -- (9,1) -- (11,1);
      \draw (0,2) -- (2,2) -- (3,1) -- (5,1) -- (6,0) -- (8,0) -- (9,0) -- (11,0);
      \draw (0,3) -- (2,3) -- (3,3) -- (5,3) -- (6,3) -- (8,3) -- (9,3) -- (11,3);
      %v
      \draw[color=red] (0,0.1) -- (0.5,0.1) -- (1.5,1.1) -- (3.5,1.1) -- (4.5,2.1) -- (6.5,2.1) -- (7.5,3.1) -- (9.5,3.1) -- (10.5,3.1) -- (11,3.1);
      \draw[color=red] (0,1.1) -- (0.5,1.1) -- (1.5,0.1) -- (3.5,0.1) -- (4.5,0.1) -- (6.5,0.1) -- (7.5,0.1) -- (9.5,0.1) -- (10.5,1.1) -- (11,1.1);
      \draw[color=red] (0,2.1) -- (0.5,2.1) -- (1.5,2.1) -- (3.5,2.1) -- (4.5,1.1) -- (6.5,1.1) -- (7.5,1.1) -- (9.5,1.1) -- (10.5,0.1) -- (11,0.1);
      \draw[color=red] (0,3.1) -- (0.5,3.1) -- (1.5,3.1) -- (3.5,3.1) -- (4.5,3.1) -- (6.5,3.1) -- (7.5,2.1) -- (9.5,2.1) -- (10.5,2.1) -- (11,2.1);
      \node[left] at (0,-0.1) {$\scriptstyle 3$};
      \node[left] at (0,0.9) {$\scriptstyle 2$};
      \node[left] at (0,1.9) {$\scriptstyle 1$};
      \node[left] at (0,2.9) {$\scriptstyle 4$};
      \node[left, color=red] at (0,0.2) {$\scriptstyle 1$};
      \node[left, color=red] at (0,1.2) {$\scriptstyle 2$};
      \node[left, color=red] at (0,2.2) {$\scriptstyle 3$};
      \node[left, color=red] at (0,3.2) {$\scriptstyle 4$};
      \node[right] at (11,-0.1) {$\scriptstyle 1$};
      \node[right] at (11,0.9) {$\scriptstyle 2$};
      \node[right] at (11,1.9) {$\scriptstyle 3$};
      \node[right] at (11,2.9) {$\scriptstyle 4$};
      \node[right, color=red] at (11,0.2) {$\scriptstyle 3$};
      \node[right, color=red] at (11,1.2) {$\scriptstyle 2$};
      \node[right, color=red] at (11,2.2) {$\scriptstyle 4$};
      \node[right, color=red] at (11,3.2) {$\scriptstyle 1$};
    \end{tikzpicture}\\
    \[\Delta_{1,3},\Delta_{2,3},\Delta_{2,1},\Delta_{3,1},\Delta_{12,23},\Delta_{12,13},\Delta_{23,13},\Delta_{23,12},\Delta_{123,123},\Delta_{234,123},\Delta_{1234,1234}\]
    \caption{A double wiring diagram for the double reduced word $(-1,2,-2,1,-3,2,-1)\in\Sigma_4$ and the collection of chamber minors determined by this double wiring diagram.}
    \label{fig:double wiring diagram}
    \end{figure}

    In a double wiring diagram, we label the strands in $u$'s wiring diagram on the right by $1$ through $n$ starting from the bottom and label the strands in $v$'s wiring diagram similarly on the left.  In this way, each chamber of the double wiring diagram determines a minor with row set the labels of strands in $v$'s wiring diagram lying below the chamber and column set the labels of strands in $u$'s wiring diagram lying below the chamber, see Figure~\ref{fig:double wiring diagram}.  Now we may describe a collection of total positivity criteria for $GL_n^{u,v}$.
    \setcounter{subtheorem}{3}
    \begin{theorem}\label{th:total positivity criteria}\cite{FZ99}
      Each double wiring diagram for $(u,v)$ determines a total positivity criterion for $GL_n^{u,v}$, i.e. an element of $GL_n^{u,v}$ lies in $GL_n^{\ge0}$ if and only if all $n+\ell(u)+\ell(v)$ chamber minors determined by the double wiring diagram are positive.
    \end{theorem}
    The totally positive matrices are exactly the elements of $GL_n^{w_0,w_0}\cap GL_n^{\ge0}$.  Thus Theorem~\ref{th:total positivity criteria} provides many total positivity criteria for $GL_n$.
    \setcounter{subexercise}{4}
    \begin{exercise}
      Use double wiring diagrams to find a total positivity criteria for $GL_3$.
    \end{exercise}
    To finish the section and connect to the theory of cluster algebras to be presented in the next section we make the following observation.  Suppose a double reduced word for $(u,v)$ contains neighboring letters, one belonging to a reduced word for $u$ and one belonging to a reduced word for $v$.  In Figure~\ref{fig:wiring exchange relation}, we show the local effect on a double wiring diagram when these letters are interchanged. 
    \begin{figure}[ht]
      \begin{tikzpicture}
        \draw (0,0) -- (1,0) -- (3,2) -- (7,2);
        \draw (0,2) -- (1,2) -- (3,0) -- (7,0);
        \draw[color=red] (0,0.1) -- (4,0.1) -- (6,2.1) -- (7,2.1);
        \draw[color=red] (0,2.1) -- (4,2.1) -- (6,0.1) -- (7,0.1);
        \node[left] at (0,-0.1) {$\scriptstyle \ell$};
        \node[left] at (0,1.9) {$\scriptstyle k$};
        \node[left, color=red] at (0,0.2) {$\scriptstyle i$};
        \node[left, color=red] at (0,2.2) {$\scriptstyle j$};
        \node[right] at (7,-0.1) {$\scriptstyle k$};
        \node[right] at (7,1.9) {$\scriptstyle \ell$};
        \node[right, color=red] at (7,0.2) {$\scriptstyle j$};
        \node[right, color=red] at (7,2.2) {$\scriptstyle i$};
        \node at (3.5,2.5) {$\scriptstyle\Delta_{I\cup\{i,j\},J\cup\{k,\ell\}}$};
        \node at (6.5,1.1) {$\scriptstyle\Delta_{I\cup\{j\},J\cup\{k\}}$};
        \node at (3.5,1.05) {$\scriptstyle\Delta_{I\cup\{i\},J\cup\{k\}}$};
        \node at (0.5,1) {$\scriptstyle\Delta_{I\cup\{i\},J\cup\{\ell\}}$};
        \node at (3.5,-0.5) {$\scriptstyle\Delta_{I,J}$};
      \end{tikzpicture}\\
      \begin{tikzpicture}
        \draw (0,0) -- (4,0) -- (6,2) -- (7,2);
        \draw (0,2) -- (4,2) -- (6,0) -- (7,0);
        \draw[color=red] (0,0.1) -- (1,0.1) -- (3,2.1) -- (7,2.1);
        \draw[color=red] (0,2.1) -- (1,2.1) -- (3,0.1) -- (7,0.1);
        \node[left] at (0,-0.1) {$\scriptstyle \ell$};
        \node[left] at (0,1.9) {$\scriptstyle k$};
        \node[left, color=red] at (0,0.2) {$\scriptstyle i$};
        \node[left, color=red] at (0,2.2) {$\scriptstyle j$};
        \node[right] at (7,-0.1) {$\scriptstyle k$};
        \node[right] at (7,1.9) {$\scriptstyle \ell$};
        \node[right, color=red] at (7,0.2) {$\scriptstyle j$};
        \node[right, color=red] at (7,2.2) {$\scriptstyle i$};
        \node at (3.5,2.5) {$\scriptstyle\Delta_{I\cup\{i,j\},J\cup\{k,\ell\}}$};
        \node at (6.5,1.1) {$\scriptstyle\Delta_{I\cup\{j\},J\cup\{k\}}$};
        \node at (3.5,1.05) {$\scriptstyle\Delta_{I\cup\{j\},J\cup\{\ell\}}$};
        \node at (0.5,1) {$\scriptstyle\Delta_{I\cup\{i\},J\cup\{\ell\}}$};
        \node at (3.5,-0.5) {$\scriptstyle\Delta_{I,J}$};
        \node at (3.5,-1) {};
      \end{tikzpicture}
      \caption{Chamber minors associated to a simple transposition of opposite letters in a double reduced word.}
      \label{fig:wiring exchange relation}
    \end{figure}

    Observe that under such an exchange we have the following exchange relation analgous to \eqref{eq:short plucker special}:
    \begin{equation}\label{eq:short Plucker}
      \Delta_{I\cup\{i\},J\cup\{k\}}\Delta_{I\cup\{j\},J\cup\{\ell\}}=\Delta_{I,J}\Delta_{I\cup\{i,j\},J\cup\{k,\ell\}}+\Delta_{I\cup\{i\},J\cup\{\ell\}}\Delta_{I\cup\{j\},J\cup\{k\}}.
    \end{equation}
    Moreover, observe that any two collections of chamber minors for double reduced words of $(u,v)$ can be obtained from each other by a sequence of these exchanges.
    \begin{exercise}
      Prove equation \eqref{eq:short Plucker}.
    \end{exercise}
  %\end{example}

  %\begin{example}
  %  <Include some example where $y$-variable mutations are natural.>
  %\end{example}

  %\begin{exercise}
  %  Anything illuminating to ask participants to do?
  %\end{exercise}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A Unifying Concept: Cluster Algebras}\label{sec:cluster_algebras}
	%We define (skew-symmetric) cluster algebras at three successive levels of generality, differing from one another in the allowable types of coefficients. 
	We now define cluster algebras of geometric type and $Y$-patterns, with a focus on the underlying dynamics of seed mutations. All definitions are due to S. Fomin and A. Zelevinsky and are drawn from \cite{FZ02} and \cite{FZ07}.
	
	\subsection{Basic definitions}
	Each cluster algebra is defined recursively from some initial data called a seed.  A seed consists of a cluster, which was informally described at the beginning of Section \ref{sec:motivation}, together with some combinatorial data that encode the mutations that can be performed.

	\begin{definition}
		A \emph{seed} is a pair $(\x,B)$ where $\x = (x_1,\ldots, x_n)$ is an $n$-tuple of elements forming a transcendence basis of a rational function field and $B$ is a skew-symmetric integer $n \times n$ matrix.  The vector $\x$ is called the \emph{cluster} and the matrix $B$ is called the \emph{exchange matrix}.
	\end{definition}
  
	The following employs the notation $[a]_+ = \max(a,0)$.
	\begin{definition}
		Given a seed ($\x,B$) and an integer $k=1,2,\ldots, n$ the \emph{seed mutation} $\mu_k$ in direction $k$ produces a new seed $\mu_k(\x,B) = (\x',B')$ where $\x' = (x_1,\ldots, x_{k-1}, x_k', x_{k+1},\ldots, x_n)$ with 
		\begin{equation} \label{eq:exchange relation}
		x_k' = \frac{\prod_{b_{ik}>0} x_i^{b_{ik}} + \prod_{b_{ik}<0} x_i^{-b_{ik}}}{x_k}
		\end{equation}
		and $B'$ is defined by
		\begin{equation}\label{eq:matrix mutation}
      b'_{ij}=\begin{cases}
                 -b_{ij} & \text{if $i=k$ or $j=k$;}\\
                 b_{ij}+[b_{ik}]_+[b_{kj}]_+-[-b_{ik}]_+[-b_{kj}]_+ & \text{otherwise.}
               \end{cases}
    \end{equation}
	\end{definition}
		
	In words, the mutation $\mu_k$ has the following effects:
	\begin{enumerate}
	\item $x_k$ changes to $x_k'$ satisfying $x_kx_k' = (\textrm{ binomial in the other } x_i$),
	\item entries $b_{ij}$ of $B$ away from row and column $k$ increase (resp. decrease) by $b_{ik}b_{kj}$ if $b_{ik}$ and $b_{kj}$ are both positive (resp. negative),
	\item row and column $k$ of $B$ are negated.
	\end{enumerate}
	
	\begin{lemma} \label{lem:mutate}
	Let $(\x,B)$ be a seed and $k=1,\ldots, n$.
	\begin{enumerate}
		\item $\mu_k(\x,B)$ is also a seed
		\item $\mu_k(\mu_k(\x,B)) = (\x,B)$.
	\end{enumerate}
	\end{lemma}
	
	\begin{definition}
		Fix an ambient field $\cF = \CC(x_1,\ldots, x_n)$ and an initial seed $((x_1,\ldots, x_n), B)$.  The entries of the clusters of all seeds reachable from this one by a sequence of mutations are called the \emph{cluster variables}.  The \emph{cluster algebra} associated with the initial seed is the subalgebra $\cA:=\cA(\bx,B)$ of $F$ generated by the set of all cluster variables.
	\end{definition}
	
  \begin{example}\label{example:type A2}
		Let $\x = (x_1,x_2)$ and
		\begin{displaymath}
			B = \left[ \begin{array}{cc}
			0 & 1 \\
			-1 & 0 \\
			\end{array}
			\right].
		\end{displaymath}
		Then $\mu_1(\x,B) = ((x_1',x_2), -B)$ where
		\begin{displaymath}
			x_1' = \frac{x_2+1}{x_1}.
		\end{displaymath}
		It is convenient to denote the new cluster variable $x_1'=x_3$.  Next $\mu_2((x_3,x_2),-B) = ((x_3,x_4),B)$ where
		\begin{displaymath}
			x_4 = \frac{x_3+1}{x_2} = \frac{\frac{x_2+1}{x_1}+1}{x_2} = \frac{x_1 + x_2 + 1}{x_1x_2}
		\end{displaymath}
		and  $\mu_2((x_3,x_4),B) = ((x_5,x_4),-B)$ where
		\begin{displaymath}
			x_5 = \frac{x_4+1}{x_3} = \frac{\frac{x_1 + x_2 + 1}{x_1x_2}+1}{\frac{x_2+1}{x_1}} = \frac{x_1+1}{x_2}.
		\end{displaymath}
		The mutation pattern seems clear, but remarkably the next variable $x_6 = (1+x_5)/x_4$ equals the first variable $x_1$.  In fact, the only distinct cluster variables that can be obtained are $x_1$ through $x_5$, so
		\begin{displaymath}
			\cA(\x,B) = \CC\left[x_1,x_2,\frac{x_2+1}{x_1},\frac{x_1 + x_2 + 1}{x_1x_2},\frac{x_1+1}{x_2}\right] \subseteq \CC[x_1^{\pm1},x_2^{\pm1}].
		\end{displaymath}
		Note the generating set is not minimal, in fact, any four of the given elements generate $\cA(\x,B)$.
    %This observation readily generalizes to cluster algebras associated to any $2\times 2$ exchange matrix. %and leads ultimately to the following important result.
  \end{example}
	
	In the preceding example, there were only finitely many cluster variables.  As we will see, this is typically not the case.  However, a more subtle feature does hold in general, namely that each of the cluster variables was a Laurent polynomial in the variables of the initial seed.  
	\begin{theorem}\label{th:Laurent phenomenon}
		For any initial seed $(\x,B)$ with $\x = (x_1,\ldots, x_n)$, the associated cluster algebra lies in the Laurent polynomial ring
		\begin{displaymath}
			\cA(\x,B) \subseteq \CC[x_1, x_1^{-1}, \ldots, x_n, x_n^{-1}].
		\end{displaymath}
		In particular, every cluster variable can be expressed as a Laurent polynomial in $x_1,\ldots, x_n$.
	\end{theorem}
	Section \ref{sec:basic_theorems} will go into more detail on this result, called the Laurent phenomenom, as well as several more main theorems on cluster algebras.
	
	\subsection{Increased generality}
	We generalize the previous definitions in two ways, first by allowing more general exchange matrices and then by allowing for certain coefficients in the exchange relations.
	
	\begin{definition}
		An $n \times n$ integer matrix $B$ is \emph{skew-symmetrizable} if there is a diagonal matrix $D$ with positive integer diagonal entries such that $DB$ is skew symmetric.
	\end{definition}
	
	As an example, a $2 \times 2$ matrix $B$ is skew-symmetrizable if and only if
	\begin{displaymath}
		B = \left[ \begin{array}{cc} 	0 & b \\	-c & 0 \\	\end{array}	\right]
		\textrm{ or } B = \left[ \begin{array}{cc} 	0 & -b \\	c & 0 \\	\end{array}	\right]
	\end{displaymath}
	for positive integers $b$ and $c$.  In either case, a possible symmetrizing matrix is
	\begin{displaymath}
		D = \left[ \begin{array}{cc} 	c & 0 \\	0 & b \\	\end{array}	\right]
	\end{displaymath}
	The \emph{rank} of a cluster algebra is the number $n$ of elements of each cluster, so the exchange matrices $B$ just described give rise to all possible rank $2$ cluster algebras.
	
	The second generalization is to allow so-called frozen variables that never mutate, but which play a part in the exchange relations for the cluster variables. An \emph{extended cluster}, by convention, is typically written $\x = (x_1,\ldots, x_n, x_{n+1},\ldots x_m)$ where $x_1,\ldots, x_n$ are the cluster variable and $x_{n+1},\ldots, x_m$ are the frozen variables.  An \emph{extended exchange matrix} is an $m \times n$ integer matrix $\tilde{B}$ with the property that its top $n \times n$ part is skew-symmetrizable.
	
	\begin{definition}
	Fix a seed $(\x, \tilde B)$ with $\x=(x_1,\ldots, x_m)$ an extended cluster and $\tilde{B}$ an extended exchange matrix.  For an integer $k = 1,2,\ldots, n$, the \emph{seed mutation} $\mu_k$ in direction $k$ produces a new seed $\mu_k(\x,\tilde{B}) = (\x',\tilde{B}')$ with $\x' = (x_1,\ldots, x_{k-1}, x_k', x_{k+1},\ldots, x_m)$.  The formulas for $x_k'$ and the entries $\tilde{b}'_{ij}$ of $B'$ are the same as in \eqref{eq:exchange relation} and \eqref{eq:matrix mutation}, where in the former the products now range from $1$ to $m$ instead of from $1$ to $n$.
	\end{definition}
	
	\begin{lemma}
	Lemma \ref{lem:mutate} holds in this generalized setting.  Moreover, the skew-symmetrizing matrix $D$ is unchanged by mutation.
	\end{lemma}
	
	Given a seed $(\x, \tilde{B})$ as above, the corresponding cluster algebra $\cA(\x, \tilde{B})$ is defined to be the subalgebra of $\CC(x_1,\ldots, x_m)$ generated by all cluster variables reachable from this seed together with the frozen variables (which appear in every seed).  Cluster algebras at this level of generality are referred to as cluster algebras of \emph{geometric type}.  In comparison, the earlier definition given was of \emph{skew-symmetric} cluster algebras with \emph{trivial coefficients}.
	
	\begin{example} \label{example:rank2}
		Consider a rank $2$ cluster algebras with trivial coefficients (so $n=2$ and $m=0$).  The initial exchange matrix can be taken to be
		\begin{displaymath}
			B = \left[ \begin{array}{cc} 	0 & b \\	-c & 0 \\	\end{array}	\right].
		\end{displaymath}
		with $b,c$ positive integers.  Then $\mu_1(B) = \mu_2(B) = -B$ and $\mu_1(-B) = \mu_2(-B) = B$, so as in Example \ref{example:type A2} (the case $b=c=1$), the cluster variables can be identified as $\{x_i : i \in \mathbb{Z}\}$ (possibly with repeats) and the mutations as 
		\begin{displaymath}
		\ldots \stackrel{\mu_1}{\longleftrightarrow} ((x_1,x_0),-B) \stackrel{\mu_2}{\longleftrightarrow} ((x_1,x_2),B)) \stackrel{\mu_1}{\longleftrightarrow} ((x_3,x_2),-B)) \stackrel{\mu_2}{\longleftrightarrow} ((x_3,x_4),B)) \stackrel{\mu_1}{\longleftrightarrow} \ldots
		\end{displaymath}
		The exchange relations are
		\begin{displaymath}
		x_{k-1}x_{k+1}=\begin{cases}x_k^b+1 & \text{if $k$ is odd;}\\ x_k^c+1 & \text{if $k$ is even.}\end{cases}
		\end{displaymath}
		
		\begin{subexercise}\label{exercise:rank 2}\mbox{}
    \begin{enumerate}
      \item Compute all cluster variables and coefficient variables for cluster algebras associated to $B=\left[\begin{array}{cc} 0 & b\\ -c & 0\end{array}\right]$ with $b,c\in\ZZ_{>0}$ and $bc\le 3$.
      \item Justify why attempting such a calculation is futile for $bc\ge4$.  Hint: consider the denominator vectors of the cluster variables.
    \end{enumerate}
  \end{subexercise}
	
	\begin{subexercise}\label{exercise:rank 2 Laurent phenomenon}
    Prove the Laurent phenomenon for rank 2 cluster algebras.  Hint: start by showing that $x_4\in\ZZ[x_0,x_1,x_2,x_3]$ by considering the monomial $x_0x_3^b$.
  \end{subexercise}
	\end{example}
	
	\begin{example}
		Let $n=3$ and $m=9$.  Denote the cluster 
		\begin{displaymath}
			\x = (\Delta_{13},\Delta_{14},\Delta_{15}, \Delta_{12},\Delta_{23},\Delta_{34},\Delta_{45},\Delta_{56},\Delta_{16})
		\end{displaymath}
		where the last $m-n=6$ variables are frozen.  Let
		\begin{displaymath}
			\tilde{B} = \left[\begin{array}{ccc}
			0 & 1 & 0 \\
			-1 & 0 & 1 \\
			0 & -1 & 0 \\
			1 & 0 & 0 \\
			-1 & 0 & 0 \\
			1 & -1 & 0 \\
			0 & 1 & -1 \\
			0 & 0 & 1 \\
			0 & 0 & -1 \\
			\end{array}
			\right].
		\end{displaymath}
		Let $\cA = \cA(\x,\tilde{B})$.  Interpret the initial variables as the indicated Plucker coordinates in $Gr_{2,6}$.
		
		\begin{subexercise} \ \newline
			\begin{itemize}
			\item The cluster variables of $\cA$ are precisely the $\Delta_{ij}$ with $1 \leq i < j \leq n$.
			\item The clusters of $\cA$ are precisely the $x_T$ (see Section \ref{sub:Gr2n}) for $T$ a triangulation of a hexagon. 
			\end{itemize}
		\end{subexercise}
		For instance, the new cluster variable produced by applying $\mu_1$ to the initial seed is
		\begin{displaymath}
		\frac{\Delta_{12}\Delta_{34} + \Delta_{14}\Delta_{23}}{\Delta_{13}}
		\end{displaymath}
		which equals $\Delta_{24}$.
	\end{example}
	
	There is an alternate formulation of cluster algebras of geometric type which focuses on the roles the frozen variables $x_{n+1},\ldots, x_m$ play in the exchange relations, rather than on the variables themselves.  Let $(\x, \tilde{B})$ be an extended seed.  For $k=1,\ldots, n$ let 
	\begin{displaymath}
	y_k = \prod_{i=n+1}^m x_i^{b_{ik}}.
	\end{displaymath}
	Define an operation (called auxiliary addition) $\oplus$ on Laurent monomials by
	\begin{displaymath}
	\prod_{i=n+1}^m x_i^{e_i} \oplus \prod_{i=n+1}^m x_i^{f_i} = \prod_{i=n+1}^m x_i^{\min(e_i,f_i)}.
	\end{displaymath}
	Using this operation, we can extract positive and negative exponents as follows
	\begin{displaymath}
	\frac{y_k}{1 \oplus y_k} = \prod_{\substack{i=n+1,\ldots, m\\ b_{ik > 0}}} x_i^{b_{ik}}
	\quad \quad \quad \frac{1}{1 \oplus y_k} = \prod_{\substack{i=n+1,\ldots, m\\ b_{ik < 0}}} x_i^{-b_{ik}}.
	\end{displaymath}
	The change in going from the original exchange relation \eqref{eq:exchange relation} to the one in geometric type, then, can be summarized by saying that the two terms of the binomial are enriched with coefficients $y_k/(1 \oplus y_k)$ and $1/(1 \oplus y_k)$.  We now write
	\begin{displaymath} 
		x_k' = \frac{y_k\prod_{b_{ik}>0} x_i^{b_{ik}} + \prod_{b_{ik}<0} x_i^{-b_{ik}}}{(1 \oplus y_k)x_k}
	\end{displaymath}
	where the products are for $i$ from $1$ to $n$.
	
	\subsection{$Y$-patterns}
	There is an alternate version of seeds and mutations, closely related to the previous, which itself arises in many applications.  
	
	\begin{definition}
		A \emph{$Y$-seed} is a pair $(\y,B)$ consisting of an $n$-tuple $\y=(y_1,\ldots, y_n)$ and an $n\times n$ skew symmetric matrix $B$.  For $k=1,\ldots, n$, the \emph{$Y$-seed mutation} $\mu_k$ is defined by
		\begin{displaymath}
			\mu_k((y_1,\ldots, y_n),B) = ((y_1',\ldots, y_n'),B')
		\end{displaymath}
		for $B'$ as defined in \eqref{eq:matrix mutation} and 
		\begin{equation}\label{eq:y mutation}
		% Warning: this seems to only agree with (2.3) in CA-IV in the skew-symmetric case
		y_j' = \begin{cases}
		y_k^{-1} & \text{if $j = k$;} \\
		y_jy_k^{|-b_{jk}|_+}(1+y_k)^{b_{jk}} & \text{if $j \neq k$.} \\
		\end{cases}
		\end{equation}
	\end{definition}
	
	In words, the $Y$-seed mutation $\mu_k$ has the following effects:
	\begin{enumerate}
	\item For each $j \neq k$, $y_j$ is multiplied by $(1+y_k)^{b_{jk}}$ if $b_{jk}>0$ and by $(1+y_k^{-1})^{b_{jk}}$ if $b_{jk}<0$,
	\item $y_k$ is inverted,
	\item $B$ is changed in the same way as for ordinary seed mutations.
	\end{enumerate}
	The $Y$-dynamics also go by the name \emph{coefficient dynamics} because the coefficients in a geometric type cluster algebra evolve in this manner with respect to auxilliary addition $\oplus$ and ordinary multiplication of Laurent monomials.
	
	Another connection between cluster algebra and $Y$-pattern dynamics comes by way of a certain Laurent monomial change of variables.  Let $\tilde{B}$ be an $m \times n$, extended exchange matrix with $B$ its upper $n \times n$ part.  Given a seed $((x_1,\ldots, x_m),\tilde{B})$, define an associated $Y$-seed $((\hat{y}_1,\ldots, \hat{y}_n),B)$ by
	\begin{displaymath}
	\hat{y}_j = \prod_{i=1}^m x_i^{b_{ij}}
	\end{displaymath}
	
	\begin{proposition}
	Fix $k=1,\ldots, n$ and suppose $\mu_k(\x,\tilde{B}) = (\x',\tilde{B}')$.  Define $(\hat{y}_1,\ldots, \hat{y}_n)$ from $(\x,\tilde{B})$ and $(\hat{y}_1',\ldots, \hat{y}_n')$ from $(\x',\tilde{B}')$ as above.  Then 
	\begin{displaymath}
	\mu_k((\hat{y}_1,\ldots, \hat{y}_n), B) = ((\hat{y}_1',\ldots, \hat{y}_n'), B').
	\end{displaymath}
	\end{proposition}

%%%%%%%%%%%%%%%%%%%%%%%
\section{Foundational Results}\label{sec:basic_theorems}
  As mentioned in the introduction cluster algebras have found applications in a surprising array of mathematical disciplines.  Much of their ubiquity comes from a number of remarkable theorems which we now explain.

  Recall the Laurent phenomenon from Theorem~\ref{th:Laurent phenomenon}.  This remarkable theorem is the prime reason for the connection between cluster algebras and many well-known classical recursions which, a priori produce rational numbers, but turn out to give integer values given certain integer inputs \cite{FZ02b}.  We will begin by explaining the solution to Exercise~\ref{exercise:rank 2 Laurent phenomenon}.  Consider the rank 2 cluster algebra with initial exchange matrix $B=\left[\begin{array}{cc}0 & b\\ -c & 0\end{array}\right]$  and cluster variables $x_k$, $k\in\ZZ$ as in Example~\ref{example:rank2}.  The cluster variables may be defined by the simple recursion
  \[x_{k-1}x_{k+1}=\begin{cases}x_k^b+1 & \text{if $k$ is odd;}\\ x_k^c+1 & \text{if $k$ is even.}\end{cases}\]
  In this case it is easy to establish the Laurent phenomenon.
  \begin{theorem}
    For any $m\in\ZZ$, each $x_k$ is an element of the ring $\ZZ[x_{m-1},x_m,x_{m+1},x_{m+2}]$.
  \end{theorem}
  \begin{remark}
    The ring $\ZZ[x_{m-1},x_m,x_{m+1},x_{m+2}]$ is called the lower bound of $\cA(\bx,B)$ at the cluster $\{x_m,x_{m+1}\}$.  We will give more details below where this result is a special case of Theorem~\ref{th:lower bounds}.
  \end{remark}
  \begin{proof}
    Without loss of generality assume $m$ is odd, otherwise interchange $b$ and $c$ in the calculation below.  Then we may compute
    \begin{align*}
      x_{m-1}x_{m+2}^b
      &=\frac{x_m^b+1}{x_{m+1}}x_{m+2}^b\\
      &=\frac{x_m^bx_{m+2}^b-1}{x_{m+1}}+\frac{x_{m+2}^b+1}{x_{m+1}}\\
      &=\frac{(x_{m+1}^c+1)^b-1}{x_{m+1}}+x_{m+3}.
    \end{align*}
    In particular, we see that $x_{m+3}\in\ZZ[x_{m-1},x_m,x_{m+1},x_{m+2}]$.  A similar calculation shows $x_{m-2}\in\ZZ[x_{m-1},x_m,x_{m+1},x_{m+2}]$ and a simple induction argument of ``shifting the viewing window'' establishes the result.
  \end{proof}
  With more work one may establish the following result which can be leveraged to prove the Laurent phenomenon in general.
  \begin{theorem}
    Assume $bc\ne0$.  For any $m\in\ZZ$, the cluster algebra $\cA(\bx,B)$ is equal to $\bigcap_{k\in\ZZ}\ZZ[x_k^{\pm1},x_{k+1}^{\pm1}]=\bigcap_{k=m-1}^{m+1}\ZZ[x_k^{\pm1},x_{k+1}^{\pm1}]$.
  \end{theorem}
  \begin{remark}
    If one allows extended exchange matrices and imposes an additional coprimality assumption (see Definition~\ref{def:coprime}, then the condition $bc\ne0$ may be dropped.  The ring $\bigcap_{k=m-1}^{m+1}\ZZ[x_k^{\pm1},x_{k+1}^{\pm1}]$ is called the upper bound of $\cA(\bx,B)$ at the cluster $\{x_m.x_{m+1}\}$ and the ring $\bigcap_{k\in\ZZ}\ZZ[x_k^{\pm1},x_{k+1}^{\pm1}]$ is called the upper cluster algebra.  We will give more details below where this result can be seen as a special case of Thoerems~\ref{th:lower bounds} and~\ref{th:acyclic upper bounds}.
  \end{remark}
  The proof of this result is somewhat technical and seems unlikely to be very informative so we will omit the details.


  As we have seen in Sections~\ref{sub:Gr2n} and~\ref{sub:total positivity}, cluster algebras have roots in classical Lie theory.  A large number of objects in this realm are classified by Dynkin diagrams and cluster algebras are no exception.  Given an extended exchange matrix $\tilde B$ with principal square submatrix $B=(b_{ij})$, write $A:=A_{\tilde B}=(a_{ij})$ for the \emph{Cartan companion} of $\tilde B$ given by 
  \[a_{ij}=\begin{cases}2 & \text{if $i=j$;}\\ -|b_{ij}| & \text{otherwise.}\end{cases}\]
  Note that the Cartan companion only depends on the principal part of $\tilde B$.
  \begin{theorem}\cite{FZ03}
    A cluster algebra $\cA(\x,\tilde B)$ has only finitely many cluster variables if and only if there exists an extended exchange matrix mutation equivalent to $\tilde B$ whose Cartan companion is a finite-type Cartan matrix.
  \end{theorem}
  \begin{example}
    As we have seen in Exercise~\ref{exercise:rank 2} the cluster algebra associated to the exchange matrix $\left[\begin{array}{cc}0 & b\\ -c & 0\end{array}\right]$ has only finitely many cluster variables if and only if $bc\le3$.  

    Writing $x_k$, $k\in\ZZ\setminus\{1,2\}$ in its initial cluster Laurent expansion, i.e. as an element of $\ZZ[x_1^{\pm1},x_2^{\pm1}]$, the exponent $d^{(k)}=(d_1^{(k)},d_2^{(k)})$ of the monomial $x_1^{d^{(k)}_1}x_2^{d^{(k)}_2}$ appearing in the denominator of $x_k$ can easily be seen to be computed as follows.  Let $U_m(t)$ denote the Chebyshev polynomial defined recursively by 
    \[U_{m+1}(t)=tU_m(t)-U_{m-1}(t)\qquad U_1(t)=1\qquad U_0(t)=0\]
    and for $j\in\{1,2\}$ define 
    \[u_{m,j}=\begin{cases}U_m(\sqrt{bc}) & \text{if $m$ is odd;}\\\sqrt{b/c}\,U_m(\sqrt{bc}) & \text{if $m$ is even and $j=1$;}\\\sqrt{c/b}\,U_m(\sqrt{bc}) & \text{if $m$ is even and $j=2$.}\\\end{cases}\]
    Note that each $u_{m,j}$ is a polynomial in $b$ and $c$.  Moreover, one may easily show that each $u_{m,j}$, $m\ge1$ is a positive integer when $bc\ge4$.  Then the denominator vector of $x_k$ for $k\in\ZZ\setminus\{1,2\}$ is
    \begin{equation}\label{eq:rank 2 denominators}
      d^{(k)}=\begin{cases}(u_{k-2,1},u_{k-3,2}) & \text{if $k\ge3$;}\\(u_{-k,1},u_{-k+1,2}) & \text{if $k\le0$.}\end{cases}
    \end{equation}
    It is not hard to check that the vectors $d^{(k)}$ are exactly the positive roots in the root system associated the the Cartan matrix $A_B=\left[\begin{array}{cc} 2 & -b\\ -c & 2\end{array}\right]$, which is known to have only finitely many roots if and only if $bc\le3$.
    \begin{subexercise}
      Prove that $U_n\big(2\cos(\theta)\big)=\frac{\sin(n\theta)}{\sin(\theta)}$ and use this to show that there are infinitely many $d^{(k)}$ if and only if $bc\ge4$.
    \end{subexercise}
  \end{example}
  Thus we see a condition on an arbitrary extended exchange matrix $\tilde B$ which implies the corresponding cluster algebra has infinitely many cluster variables: if there exists an exchange matrix $\tilde B'=(b'_{ij})$ mutation equivalent to $\tilde B$ for which $b'_{ij}b'_{ji}\ge4$ for some $i,j$, then the cluster algebra $\cA(\bx,\tilde B)$ is not of finite-type.  It is quite remarkable that the opposite condition is enough to characterize finite-type cluster algebras.  Call an extended exchange matrix $\tilde B=(b_{ij})$ \emph{2-finite} if every matrix $\tilde B'=(b'_{ij})$ mutation equivalent to $\tilde B$ satisfies the rank 2 finite-type condition $b'_{ij}b'_{ji}\le3$ for all $i,j$.
  \begin{theorem}\cite{FZ03}
    A cluster algebra $\cA(\x,\tilde B)$ has only finitely many cluster variables if and only if $\tilde B$ is 2-finite.
  \end{theorem}

  The next result gives the positivity of the initial cluster Laurent expansions of all cluster variables.
  \begin{theorem}\cite{LS15}
    Let $\cA(\bx,\tilde B)$ be a cluster algebra.  Any cluster variable of $\cA(\bx,\tilde B)$ is an element of $\ZZ_{\ge0}[x_1^{\pm1},\ldots,x_m^{\pm1}]$.
  \end{theorem}
  The proof builds on a concrete combinatorial construction of cluster variables in rank 2 \cite{LS13,LLZ14}.  To describe it we need more notation.

  For $k\in\ZZ\setminus\{1,2\}$, denote by $R_k$ the rectangle in $\ZZ^2$ with corner vertices $(0,0)$ and $d^{(k)}$ from Equation~\ref{eq:rank 2 denominators}.  Write $D_k$ for the maximal Dyck path in $R_k$ beginning at $(0,0)$, taking East and North steps to end at the upper right corner of $R_k$ while never passing above the main diagonal, and such that the area below $D_k$ inside $R_k$ is maximized.
  \begin{example}
    For $b=c=3$, the maximal Dyck path $D_5$ is
    \begin{center}
    \begin{tikzpicture}
      \draw[step=0.5cm,color=gray] (0,0) grid (4.0,1.5);
      \draw[color=gray] (0,0) -- (4.0,1.5);
      \draw[fill=black] (0.0,0.0) circle (2.2pt);
      \draw[fill=black] (0.5,0.0) circle (2.2pt);
      \draw[fill=black] (1.0,0.0) circle (2.2pt);
      \draw[fill=black] (1.5,0.0) circle (2.2pt);
      \draw[fill=black] (1.5,0.5) circle (2.2pt);
      \draw[fill=black] (2.0,0.5) circle (2.2pt);
      \draw[fill=black] (2.5,0.5) circle (2.2pt);
      \draw[fill=black] (3.0,0.5) circle (2.2pt);
      \draw[fill=black] (3.0,1.0) circle (2.2pt);
      \draw[fill=black] (3.5,1.0) circle (2.2pt);
      \draw[fill=black] (4.0,1.0) circle (2.2pt);
      \draw[fill=black] (4.0,1.5) circle (2.2pt);
    \end{tikzpicture}
    \end{center}
  \end{example}
  For edges $e,e'\in D_k$, write $e<e'$ if $e$ precedes $e'$ along $D_k$, in other words if $e$ is closer to $(0,0)$.  In this case write $ee'$ for the subpath of $D_k$ beginning with the edge $e$ and ending with the edge $e'$.  Write $(ee')_H$ and $(ee')_V$ for the sets of horizontal and vertical edges in the path $ee'$ respectively.  Let $H,V\subset D_k$ denote the sets of horizontal and vertical edges of $D_k$.

  Call subsets $S_H\subset H$ and $S_V\subset V$ compatible if for each $h\in S_H$ and $v\in S_V$ with $h<v$, there exists $e\in hv$ so that one of the following holds:
  \begin{itemize}
    \item $e\ne v$ and $|(he)_V|=c|(he)_H\cap S_H|$;
    \item $e\ne h$ and $|(ev)_H|=b|(ev)_V\cap S_V|$.
  \end{itemize}
  \begin{example}
    For $b=c=3$, the picture below on the left shows a compatible collection of edges of $D_5$ while the collection on the right is not compatible.
    \begin{center}
    \begin{tikzpicture}
      \draw[step=0.5cm,color=gray] (0,0) grid (4.0,1.5);
      \draw[color=gray] (0,0) -- (4.0,1.5);
      \draw[fill=black] (0.0,0.0) circle (2.2pt);
      \draw[fill=black] (0.5,0.0) circle (2.2pt);
      \draw[line width=2.5pt] (0.5,0.0) -- (1.0,0.0);
      \draw[fill=black] (1.0,0.0) circle (2.2pt);
      \draw[fill=black] (1.5,0.0) circle (2.2pt);
      \draw[fill=black] (1.5,0.5) circle (2.2pt);
      \draw[fill=black] (2.0,0.5) circle (2.2pt);
      \draw[fill=black] (2.5,0.5) circle (2.2pt);
      \draw[fill=black] (3.0,0.5) circle (2.2pt);
      \draw[line width=2.5pt] (3.0,0.5) -- (3.0,1.0);
      \draw[fill=black] (3.0,1.0) circle (2.2pt);
      \draw[fill=black] (3.5,1.0) circle (2.2pt);
      \draw[fill=black] (4.0,1.0) circle (2.2pt);
      \draw[line width=2.5pt] (4.0,1.0) -- (4.0,1.5);
      \draw[fill=black] (4.0,1.5) circle (2.2pt);
    \end{tikzpicture}
    \qquad
    \begin{tikzpicture}
      \draw[step=0.5cm,color=gray] (0,0) grid (4.0,1.5);
      \draw[color=gray] (0,0) -- (4.0,1.5);
      \draw[fill=black] (0.0,0.0) circle (2.2pt);
      \draw[fill=black] (0.5,0.0) circle (2.2pt);
      \draw[fill=black] (1.0,0.0) circle (2.2pt);
      \draw[line width=2.5pt] (1.0,0.0) -- (1.5,0.0);
      \draw[fill=black] (1.5,0.0) circle (2.2pt);
      \draw[fill=black] (1.5,0.5) circle (2.2pt);
      \draw[fill=black] (2.0,0.5) circle (2.2pt);
      \draw[fill=black] (2.5,0.5) circle (2.2pt);
      \draw[fill=black] (3.0,0.5) circle (2.2pt);
      \draw[line width=2.5pt] (3.0,0.5) -- (3.0,1.0);
      \draw[fill=black] (3.0,1.0) circle (2.2pt);
      \draw[fill=black] (3.5,1.0) circle (2.2pt);
      \draw[fill=black] (4.0,1.0) circle (2.2pt);
      \draw[line width=2.5pt] (4.0,1.0) -- (4.0,1.5);
      \draw[fill=black] (4.0,1.5) circle (2.2pt);
    \end{tikzpicture}
    \end{center}
    \begin{subexercise}
      Find all other compatible collections for $D_5$ in the case $b=c=3$.
    \end{subexercise}
  \end{example}
  \begin{theorem}\label{th:rank 2 positivity}\cite{LLZ14}
    For $k\in\ZZ\setminus\{1,2\}$, the cluster variable $x_k$ is given by
    \[x_k=\sum\limits_{(S_H,S_V)}x_1^{-d^{(k)}_1+b|S_V|}x_2^{-d^{(k)}_2+c|S_H|},\]
    where the sum ranges over all compatible collections of edges in the maximal Dyck path $D_k$.  In particular, each cluster variable is contained in $\ZZ_{\ge0}[x_1^{\pm1},x_2^{\pm1}]$.
  \end{theorem}

  \begin{remark}
    The idea of obtaining positivity for arbitrary cluster algebras goes as follows (see \cite{LS15}): any sequence of mutations can be viewed as a collection of rank 2 mutation sequences, so iteratively applying Theorem~\ref{th:rank 2 positivity} will (after a considerable amount of work) lead to a proof of positivity in general.
  \end{remark}

  One motivation for the discovery of the cluster algebra formalism was the desire to find a combinatorial construction of dual canonical basis elements for (quantum) algebraic groups.  The dual canonical basis of a semisimple algebraic group induces bases on the coordinate rings of any of its subvarieties.  As we saw in Example~\ref{sub:total positivity} there appears to be some kind of cluster algebra structure on the double Bruhat cells of a semisimple algebraic group, which one can hope will shed some light on the dual canonical basis.  In most cases the double Bruhat cell does not actually admit a cluster algebra structure but does admit a closely related structure.

  \begin{definition}
    Let $(\x,\tilde B)$ denote an extended seed.  Define the \emph{upper cluster algebra} $\overline{\cA}(\x,\tilde B)$ as the intersection of all Laurent rings associated to extended seeds mutation equivalent to $(\x,\tilde B)$.
  \end{definition}

  \begin{remark}
    The upper cluster algebra is exactly the collection of all rational functions in $\{x_1,\ldots,x_m\}$ which are expressed as Laurent polynomials in terms of any cluster.  Thus the Laurent phenomenon Theorem~\ref{th:Laurent phenomenon} establishes the inclusion $\cA(\x,\tilde B)\subset\overline{\cA}(\x,\tilde B)$, hence the name ``upper'' cluster algebra.
  \end{remark}

  \begin{theorem}\label{th:dbc is uca}\cite{BFZ05}
    The coordinate ring of any double Bruhat cell $GL_n^{u,v}$ is an upper cluster algebra such that each collection of chamber minors determines a cluster.
  \end{theorem}

  In some special cases, the cluster algebra coincides with its upper cluster algebra.  First we describe a necessary intermediate concept.
  \begin{definition}\label{def:coprime}
    Let $(\bx,\tilde B)$ be an extended seed and denote by $P_k:=x_kx_k'$ the binomial on the right hand side of the exchange relation \eqref{eq:exchange relation}.  We say that $(\bx,\tilde B)$ is \emph{coprime} if the binomials $P_k$ are pairwise relatively prime.
  \end{definition}
  The \emph{upper bound} $\cU(\bx',\tilde B)$ for $\cA(\bx,\tilde B)$ at a seed $(\bx',\tilde B')$ mutation equivalent to $(\bx,\tilde B)$ is the intersection of the Laurent ring generated by variables in $\bx'$ with all Laurent rings associated to seeds which can be obtained from $(\bx',\tilde B')$ by a single mutation.
  \begin{theorem}\cite{BFZ05}
    If $(\bx,\tilde B)$ and $(\bx',\tilde B')$ are mutation equivalent coprime seeds, then their upper bounds $\cU(\bx,\tilde B)$ and $\cU(\bx',\tilde B)$ coincide.  In particular, if every seed mutation equivalent to an extended seed $(\bx,\tilde B)$ is coprime, then $\overline{\cA}(\bx,\tilde B)=\cU(\bx',\tilde B')$ for any seed mutation equivalent to $(\bx,\tilde B)$.
  \end{theorem}

  \begin{definition}
    An exchange matrix $B=(b_{ij})$ is called \emph{acyclic} if there exists a permutation $\sigma$ so that $b_{\sigma_i\sigma_j}\le 0$ for $i<j$.
  \end{definition}
  This terminology can be easily understood in the case of a skew-symmetric exchange matrix, it is acyclic exactly when its associated quiver has no oriented cycles.
  \begin{theorem}\cite{BFZ05}\label{th:acyclic upper bounds}
    Let $\cA(\x,\tilde B)$ be a cluster algebra where $\tilde B$ is coprime and has acyclic principal part.  Then the cluster algebra $\cA(\x,\tilde B)$ coincides with its upper bound $\cU(\x,\tilde B)$.
  \end{theorem}

  Acyclicity also guarantees the existence of easily identifiable bases of a cluster algebra.  These are best understood by identifying the cluster algebra with another related algebra, its lower bound.
  \begin{definition}
    Let $(\x,\tilde B)$ be an extended seed and write $x'_k$ for the variable obtained by mutation in direction $k$.  The \emph{lower bound} of $\cA(\x,\tilde B)$ at $(\x,\tilde B)$ is the subalgebra $\cL(\x,\tilde B)=\CC[x_1,x'_1,\ldots,x_n,x'_n,x_{n+1},\ldots,x_m]$.
  \end{definition}

  \begin{theorem}\cite{BFZ05}\label{th:lower bounds}
    Let $\cA(\x,\tilde B)$ be a cluster algebra where $\tilde B$ is coprime and has acyclic principal part.  Then the cluster algebra $\cA(\x,\tilde B)$ coincides with its lower bound $\cL(\x,\tilde B)$.  Moreover, the collection of all monomials in $x_1,x'_1,\ldots,x_n,x'_n,x_{n+1},\ldots,x_m$ which do not contain both $x_k$ and $x_k'$ for any $k$ forms a basis of the cluster algebra $\cA(\bx,\tilde B)$.
  \end{theorem}

  \begin{example}
    Acyclic cluster algebras are particularly nice but should not be taken as the general case, there exist cluster algebras where anything that can go wrong seems to go wrong.  The primary example of this is the \emph{Markov cluster algebra} with exchange matrix 
    \[B=\left[\begin{array}{ccc} 0 & 2 & -2\\ -2 & 0 & 2\\ 2 & -2 & 0\end{array}\right].\]
    
    Comparing with Example~\ref{sub:Markov triples} we see that the Markov cluster algebra has infinitely many cluster variables.  In this case, the following undesirable properties all hold:
    \begin{itemize}
      \item the standard monomials are linearly dependent;
      \item the Markov cluster algebra is not finitely generated and not Noetherian;
      \item the Markov cluster algebra (over $\CC$) contains non-prime cluster variables;
      \item the Markov cluster algebra does not equal its upper cluster algebra.
    \end{itemize}
  \end{example}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Compatible Poisson Structures and Quantization}\label{sec:poisson_and_quantum}

  Many natural examples of cluster algebras carry the following additional structure. 
  \begin{definition}
    A \emph{Poisson algebra} $(A,\{\cdot,\cdot\})$ is an associative algebra $A$ equipped with an additional skew-symmetric bilinear operation, called the Poisson bracket (written as $\{x,y\}$ for $x,y\in A$), for which the following holds:
    \begin{itemize}
      \item given any $x\in A$, the automorphism $\{x,\cdot\}:A\to A$, $y\mapsto\{x,y\}$ is a derivation with respect to both binary operations on $A$, i.e. for $x,y,z\in A$ we have
      \begin{align*}
        \tag{Leibnitz rule} \{x,yz\}&=\{x,y\}z+y\{x,z\}\\
        \tag{Jacobi identity}\{x,\{y,z\}\}&=\{\{x,y\},z\}+\{y,\{x,z\}\}.
      \end{align*}
    \end{itemize}
  \end{definition}

  In each such example the Poisson brackets of cluster variables take an especially simple form.  
  \begin{definition}
    Let $\cA$ be a cluster algebra.  A Poisson bracket $\{\cdot,\cdot\}$ on $\cA$ is compatible with the cluster algebra structure if every cluster $\bx$ of $\cA$ is \emph{log-canonical} with respect to $\{\cdot,\cdot\}$, that is $\{x_i,x_j\}=\Omega_{ij}x_ix_j$ for some skew-symmetric matrix $\Omega_\bx$ and all $i,j$.
  \end{definition}

  The existence of a compatible Poisson structure is quite restrictive and requires a particular relationship between the matrix $\Omega_\bx$ and the exchange matrix of a seed with cluster $\bx$. 
  \begin{lemma}
  \label{le:compatibility}
    Let $\cA(\bx,B)$ be a cluster algebra with compatible Poisson structure $\{\cdot,\cdot\}$.  Then writing $\Omega_\bx$ for the matrix of Poisson bracket coefficients, we have $B^T\Omega_\bx=[D\, \boldsymbol{0}]$, where $D$ is a diagonal matrix which skew-symmetrizes $B$.
  \end{lemma}
  \begin{subexercise}
    Prove Lemma~\ref{le:compatibility}. 
  \end{subexercise}
  
  In fact, these conditions can only hold in the following situation, in which case all possible compatible Poisson structures can be easily understood. 
  \begin{theorem}\cite{GSV03} \label{thm:Poisson}
    A compatible Poisson structure exists if and only if the extended exchange matrix $\tilde B$ has full rank.  Moreover, the collection of all such Poisson structures is parametrized by an affine space of dimension $\rho(B)+{m\choose 2}$, where $\rho(B)$ denotes the number of connected components of $B$.
  \end{theorem}
  The primary motivating example for compatible Poisson structures comes in the form of the double Bruhat cells of Section~\ref{sub:total positivity}.  To introduce the relevant Poisson structure and place the double Bruhat cells in the proper context we need to introduce more notation.

  A smooth manifold $M$ together with a Poisson structure on its algebra of $\cO(M)$ regular functions is called a \emph{Poisson manifold}.  The best understood examples of such structures comes from the theory of symplectic geometry.
  \begin{example}
    A symplectic manifold $(M,\omega)$ is a smooth even-dimensional manifold $M^{2n}$ together with a non-degenerate 2-form $\omega\in H^2(M)$, i.e. $\omega^n\ne0$ is a volume form.  The symplectic structure $\omega$ provides a natural association of a vector field $\xi_f$ to a function $f\in\cO(M)$ via the contraction formula $\iota_{\xi_f}\omega=-df$.  The algebra of smooth functions on the symplectic manifold $M$ is then naturally a Poisson algebra via $\{f,g\}=\omega(\xi_f,\xi_g)$ for $f,g\in\cO(M)$.
  \end{example}  
  A smooth map $\varphi:M\to N$ between Poisson manifolds is a \emph{Poisson morphism} if it induces a Poisson morphism $\varphi^*:\cO(N)\to\cO(M)$ on their algebras of smooth functions, i.e. $\{\varphi^*(f),\varphi^*(g)\}_M=\varphi^*\big(\{f,g\}_N\big)$ for all $f,g\in \cO(N)$.  Given two Poisson manifolds $M$ and $N$, there is a natural Poisson structure on $M\times N$ given by
  \[\{f,g\}(x,y)=\{f(\cdot,y),g(\cdot,y)\}_M(x)+\{f(x,\cdot),g(x,\cdot)\}_N(y)\]
  for $f,g\in\cO(M\times N)$.
  \begin{example}
    The general linear group $GL_n$ is a Poisson manifold where the Poisson bracket on matrix entries is given by
    \begin{equation}
      \{x_{ij},x_{k\ell}\}=\frac{1}{2}\big(\sgn(k-i)+\sgn(\ell-j)\big)x_{i\ell}x_{kj}.
    \end{equation}
    In fact, $GL_n$ has the structure of a Poisson-Lie group, that is the multiplication map $GL_n\times GL_n\to GL_n$ is a Poisson morphism when $GL_n\times GL_n$ is given the product Poisson structure.
  \end{example}
  With this we may describe the Poisson structure on the double Bruhat cells.
  \begin{theorem}\cite{GSV10}
    The cluster algebra structure on $GL_n^{u,v}$ from Theorem~\ref{th:dbc is uca} is compatible with the restriction of the Poisson structure from $GL_n$.
  \end{theorem}

  The double Bruhat cells actually have a much closer connection to the Poisson geometry of $GL_n$.  To describe it we need to introduce a few additional concepts.

  The existence of a Poisson structure on a manifold allows one to construct vector fields associated to functions in a similar manner to the symplectic case.
  \begin{lemma}\label{le:Hamiltonian vector fields}
    Let $(M,\{\cdot,\cdot\})$ be a Poisson manifold.  For any $h\in\cO(M)$, there exists a vector field $\xi_h$ on $M$ such that for any $f\in\cO(M)$ we have
    \[\xi_h(f)=\{f,h\}=\langle df,\xi_h\rangle,\]
    where $\langle\cdot,\cdot\rangle$ denotes the standard pairing between cotangent and tangent vectors over a point of $M$.
  \end{lemma}
  The vector field $\xi_h$ from Lemma~\ref{le:Hamiltonian vector fields} is simply the vector field canonically associated to the derivation $\{\cdot,h\}$ and is called a \emph{Hamiltonian vector field}.  Flowing along Hamiltonian vector fields defines an equivalence relation on a Poisson manifold as follows: 
  \begin{quotation}
    $p\sim q$ if there exists a piecewise smooth curve connecting $p$ and $q$ where each smooth segment is the trajectory of a Hamiltonian vector field.
  \end{quotation}
  The equivalence class $M_p$ containing $p\in M$ is a connected submanifold of $M$ and by definition its tangent space $T_qM_p\subset T_qM$ is spanned by Hamiltonian tangent vectors.  We may view $M_p$ as a symplectic manifold whose 2-form is given by $\omega(\xi_f,\xi_g)=\{f,g\}$ and thus $M_p$ is called the \emph{symplectic leaf} of $M$ through $p$.
  \begin{example}
    The symplectic leaves of $GL_n$ have been explicitly computed (e.g. see \cite{KZ02} and references therein).  In particular, it is known that each double Bruhat cell in $GL_n$ is foliated by symplectic leaves of $GL_n$ and forms an orbit under the action of a torus.  As an example (c.f. Example~\ref{example:big cell}) the following set is a symplectic leaf living in $GL_n^{w_0,w_0}$:
    \[\{M\in GL_n:\Delta_{[1,i],[n+1-i,n]}(M)=\Delta_{[i+1,n],[1,n-i]}(M)\ne0\text{ for all $i$}\}.\]
  \end{example}
  \begin{exercise}
    Find a matrix $M\in GL_2$ so that the symplectic leaf above is the symplectic leaf of $GL_2$ through $M$.  Do the same for $GL_3$.
  \end{exercise}

  A final reason to care about Poisson structures on a cluster algebra is that they provide a canonical quantization of the cluster algebra structure.
  \begin{theorem}\cite{berenstein-zelevinsky}
    
  \end{theorem}
  Given a cluster algebra $\cA(\bx,\tilde B)$ of full rank, any choice of compatible Poisson structure gives rise to a canonical quantization of $\cA(\bx,\tilde B)$ by the following procedure where $v$ denotes a formal variable:
  \begin{itemize}
    \item Each cluster should be replaced by a collection of quasi-commuting elements $\{X_1,\ldots,X_m\}$ satisfying $X_iX_j=v^{2\Lambda(\varepsilon_j,\varepsilon_i)}X_jX_i$ for a skew-symmetric bilinear form $\Lambda:\ZZ^m\times\ZZ^m\to\ZZ$.  That is, this \emph{quantum cluster} generates a \emph{quantum torus} $\cT=\ZZ[v^{\pm1}]\langle X_1^{\pm1},\ldots,X_m^{\pm1}\rangle$ with a $\ZZ[v^{\pm1}]$-basis 
    \[\left\{X^\ba:=v^{\sum\limits_{i<j}a_ia_j\Lambda(\varepsilon_i,\varepsilon_j)}X_1^{a_1}\cdots X_m^{a_m}\Big|\ba=(a_1,\ldots,a_m)\in\ZZ^m\right\}.\]
    \item Each quantum cluster obtained by mutation from this one should again generate a quantum torus, this forces a compatibility condition on the exchange matrix $\tilde B$ and the commutation matrix $\Lambda$ (by a slight abuse of notation we freely identify the bilinear form $\Lambda$ with a skew-symmetric matrix), that is 
    \begin{equation}\label{eq:quantum compatibility}
      \Lambda(\bb_k,\varepsilon_\ell)=0\quad\text{if $k\ne\ell$,}
    \end{equation}
    where $\bb_k$ denotes the $k^{th}$ column of $\tilde B$ thought of as an element of $\ZZ^m$.
    \item The above conditions still leave a great deal of freedom.  To restrict further, notice that the quasi-commutation relation above can be rewritten as $v^{\Lambda(\varepsilon_i,\varepsilon_j)}X_iX_j=v^{\Lambda(\varepsilon_j,\varepsilon_i)}X_jX_i$, in particular this implies that $\cT$ admits an anti-automorphism $X\mapsto \overline{X}$ which fixes each $X_i$ and sends $v$ to $v^{-1}$ (recall that $\Lambda$ was assumed skew-symmetric).  Since each initial cluster variable was \emph{bar-invariant}, the same should be true of all cluster variables since a cluster algebra is independent of the choice of initial cluster.  This uniquely determines how to incorporate powers of $v$ into the exchange relations \eqref{eq:exchange relation}, more precisely the new variable obtained by mutation in direction $k$ must be
    \[X_k'=X^{-\varepsilon_k+\sum\limits_{b_{ik}>0}b_{ik}\varepsilon_i}+X^{-\varepsilon_k-\sum\limits_{b_{ik}<0}b_{ik}\varepsilon_i}.\]
  \end{itemize}
  \begin{exercise}
    Verify the compatibility condition \eqref{eq:quantum compatibility} and show that it is the same as the compatibility condition from Lemma~\ref{le:compatibility}.
  \end{exercise}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Applications to Integrable Systems}\label{sec:integrable_systems}
	\begin{definition}
		A \emph{quiver} is a directed graph without loops (arrows from a vertex to itself) and without oriented $2$-cycles.
	\end{definition}
	
	A skew-symmetric $n\times n$ exchange matrix $B$ is often represented instead as a quiver $Q$ on vertex set $\{1,2,\ldots, n\}$.  More precisely, if $i,j$ are such that $b_{ij}>0$ then $Q$ has $b_{ij}$ arrows from $i$ to $j$ (and none from $j$ to $i$).  Matrix mutation on $B$ induces so-called quiver mutation on $Q$.
	
	\begin{definition}
		Given some $k=1,2,\ldots, n$, the \emph{quiver mutation} $\mu_k$ gives rise to $Q' = \mu_k(Q)$ by applying the following steps to $Q$:
		\begin{enumerate}
		\item for each pair of vertices $i,j$ and arrows $i\to k$ and $k \to j$, add an arrow $i \to j$,
		\item reverse all arrows $i \to k$ and $k \to j$,
		\item erase in turn pairs of arrows $i \to j$ and $j \to i$ to eliminate any $2$-cycles.
		\end{enumerate}
	\end{definition}
	
	Given a quiver $Q$ and a sequence $k_1,k_2,\ldots, k_l$ of its vertices (possibly with repeats), one can consider the dynamical system obtained by repeated applications of the composite mutation $\mu_{k_l} \circ \cdots \circ \mu_{k_2} \circ \mu_{k_1}$ starting from an initial seed $(\x, Q)$ or $Y$-seed $(\y,Q)$.  In the special case when the final quiver after one application equals the initial one, i.e.
	\begin{displaymath}
		\mu_{k_l} \circ \cdots \circ \mu_{k_2} \circ \mu_{k_1}(Q) = Q,
	\end{displaymath}
	the system amounts to iterating a fixed birational transformation.  It is natural to investigate these mappings for properties such as periodicity and integrability.  We focus on two (families of) quivers, one giving rise to the $Y$-system of type $A \times A$ and the other corresponding to the pentagram map. Some of the many other systems that have been investigated from this point of view are $Q$-systems and $T$-systems, Somos sequences, and mutations in box products of quivers. 

For the remainder of this section, all mutations are understood to be of $Y$-pattern type.

\subsection{Zamolodchikov periodicity}
	Fix positive integers $r$ and $s$ and consider the quiver $Q$ on vertex set $\{1,\ldots, r \} \times \{1,\ldots, s\}$ as illustrated in the case $r=4$, $s=3$ by
	
	\begin{pspicture}(5,4)
		\cnode(1,1){0.1}{v11}
		\cnode(2,1){0.1}{v21}
		\cnode(3,1){0.1}{v31}
		\cnode(4,1){0.1}{v41}
		\cnode(1,2){0.1}{v12}
		\cnode(2,2){0.1}{v22}
		\cnode(3,2){0.1}{v32}
		\cnode(4,2){0.1}{v42}
		\cnode(1,3){0.1}{v13}
		\cnode(2,3){0.1}{v23}
		\cnode(3,3){0.1}{v33}
		\cnode(4,3){0.1}{v43}
		\ncline{->}{v11}{v21}
		\ncline{->}{v31}{v21}
		\ncline{->}{v31}{v41}
		\ncline{<-}{v12}{v22}
		\ncline{<-}{v32}{v22}
		\ncline{<-}{v32}{v42}
		\ncline{->}{v13}{v23}
		\ncline{->}{v33}{v23}
		\ncline{->}{v33}{v43}
		\ncline{<-}{v11}{v12}
		\ncline{<-}{v13}{v12}
		\ncline{->}{v21}{v22}
		\ncline{->}{v23}{v22}
		\ncline{<-}{v31}{v32}
		\ncline{<-}{v33}{v32}
		\ncline{->}{v41}{v42}
		\ncline{->}{v43}{v42}
	\end{pspicture}
	
	Let $\mu_{\textrm{even}}$ be the compound mutation given by applying each $\mu_{i,j}$ with $i+j$ even in turn.  As no arrows connect these even vertices, it follows that these mutations commute so the order does not matter.  Define $\mu_{\textrm{odd}}$ similarly as the compound mutation at all odd vertices.
	
	\begin{exercise}
		Let $Y = (Y_{i,j})_{i=1,\ldots, r; j=1,\ldots, s}$.  Then $\mu_{\textrm{even}}(Y,Q) = (Y',-Q)$ where
		\begin{equation} \label{eq:YSystemCases}
			Y_{i,j}' = \begin{cases}
			Y_{i,j}^{-1} & \textrm{if $i+j$ even} \\
			Y_{i,j}\frac{\displaystyle\prod_{|i-i'|=1}(1+Y_{i',j})}{\displaystyle\prod_{|j-j'|=1}(1+Y_{i,j'}^{-1})} & \textrm{if $i+j$ odd} \\
			\end{cases}
		\end{equation}
		and $-Q$ denotes $Q$ with all its arrows reversed.
	\end{exercise}
	
	Now begin with the $Y$-seed $(Y_0,Q)$ and let 
	\begin{align*}
	(Y_1,-Q) &= \mu_{\textrm{even}}(Y_0,Q) \\ 
	(Y_2,Q) &= \mu_{\textrm{odd}}(Y_1,-Q) \\ 
	(Y_3,-Q) &= \mu_{\textrm{even}}(Y_2,Q) \\ 
	&\vdots
	\end{align*}
	where $Y_t = (Y_{ijt})_{i=1,\ldots, r; j=1,\ldots, s}$.  It follows from the above that
	\begin{displaymath}
		Y_{i,j,t+1} = \begin{cases}
		Y_{i,j,t}^{-1} & \textrm{if $i+j+t$ even} \\
		Y_{i,j,t}\frac{\displaystyle\prod_{|i-i'|=1}(1+Y_{i',j,t})}{\displaystyle\prod_{|j-j'|=1}(1+Y_{i,j',t}^{-1})} & \textrm{if $i+j+t$ odd} \\
		\end{cases}
	\end{displaymath}
	It is convenient to consider only the $Y_{i,j,t}$ with $i+j+t$ even, which satisfy their own recurrence
	\begin{equation} \label{eq:YSystem}
		Y_{i,j,t-1}Y_{i,j,t+1} = \frac{\displaystyle\prod_{|i-i'|=1}(1+Y_{i',j,t})}{\displaystyle\prod_{|j-j'|=1}(1+Y_{i,j',t}^{-1})}
	\end{equation}
	for all $i+j+t$ odd.  This recurrence is called the type $A_r \times A_s$ $Y$-system, a special case of a family of systems conjectured by Zamolodchikov to be periodic.  The proof, in this case, is due to A. Volkov \cite{V07}.
	
	\begin{theorem}
		The $Y$-system \eqref{eq:YSystem} on variables $Y_{i,j,t}$ with $i+j+t$ odd, $i=1,\ldots, r$, and $j=1,\ldots, s$ has period $2(r+s+2)$ in the $t$-direction, that is
		\begin{displaymath}
			Y_{i,j,t+2(r+s+2)} = Y_{i,j,t}.
		\end{displaymath}
	\end{theorem}
	
	Returning to the $Y$-pattern point of view, the initial seed $(Y_0,Q)$ consists of the $Y_{i,j,0}$ for $i+j$ even and the $Y_{i,j,-1}^{-1}$ for $i+j$ odd.  The seed $(Y_2,Q) = \mu_{\textrm{odd}}(\mu_{\textrm{even}}(Y_0,Q))$ consists of the $Y_{i,j,2}$ for $i+j$ even and the $Y_{i,j,1}^{-1}$ for $i+j$ odd.  The periodicity theorem, then, asserts that the rational map $\mu_{\textrm{odd}} \circ \mu_{\textrm{even}}$ has order $r+s+2$.
	
	\subsection{The pentagram map} The pentagram map is a discrete dynamical system defined on the space of polygons in the projective plane.  Figure \ref{fig:pentagram} shows a polygon $A$ and the corresponding output $B=T(A)$, with $T$ being the notation for the pentagram map.  Each vertex of $B$ lies at the intersection of two consecutive ``shortest'' diagonals of $A$.
	
	\begin{figure} \label{fig:pentagram}
	\begin{pspicture}(5,5)
	\pspolygon[linewidth=2pt](1,2)(1,3)(2,4.5)(3,5)(4.5,4.5)(5,3.5)(4.5,2)(3.5,1)(2,1)
  \pspolygon[linestyle=dashed](1,2)(2,4.5)(4.5,4.5)(4.5,2)(2,1)(1,3)(3,5)(5,3.5)(3.5,1)
  \pspolygon[linewidth=2pt](1.22,2.55)(1.67,3.67)(2.5,4.5)(3.67,4.5)(4.5,3.87)(4.5,2.67)(3.97,1.79)(2.75,1.3)(1.62,1.75)
	\uput[ul](2,4.5){$A$}
	\uput[dr](1.67,3.67){$B=T(A)$}
	\end{pspicture}
	\caption{An application of the pentagram map}
	\end{figure}
	
	The pentagram map was introduced by R. Schwartz \cite{S92} and extensively studied by V. Ovsienko, Schwartz and S. Tabachnikov \cite{OST10, OST13}.  In particular, they demonstrated that the pentagram map is a complete integrable system.  Another take on integrability was provided by F. Soloviev \cite{S13}.  In the same span of time, M. Glick \cite{G11} described the pentagram map as mutations in a $Y$-pattern, building on work of Schwartz \cite{S08} who had found a lift to the octahedron recurrence.  Finally, M. Gekhtman, M. Shapiro, Tabachnikov and A. Vainshtein \cite{GSTV12} gave a uniform treatment of integrability and ``clusterness'' of the pentagram map, and generalizations thereof, in terms of weighted networks on tori.  A similar framework is provided by the cluster integrable systems of A. Goncharov and R. Kenyon \cite{GK13}, albeit without any explicit mention of the pentagram map.
	
	The connection between the pentagram map and $Y$-patterns comes by way of certain geometrically defined coordinates on the space of polygons.  The \emph{cross ratio} of $4$ real numbers $x_1,x_2,x_3,x_4$ is 
	\begin{displaymath}
		\chi(x_1,x_2,x_3,x_4) = \frac{(x_1-x_2)(x_3-x_4)}{(x_1-x_3)(x_2-x_4)}.
	\end{displaymath}
	The cross ratio is invariant under projective transformations, that is 
	\begin{displaymath}
		\chi(f(x_1),f(x_2),f(x_3),f(x_4)) = \chi(x_1,x_2,x_3,x_4)
	\end{displaymath}
	for any fractional linear map $f(x) = (ax+b)/(cx+d)$.  For this reason, there is a well-defined notion in the plane of the cross ratio of 4 collinear points or of 4 concurrent lines.
	
	Given an $n$-gon $A$, label its sides and vertices consecutively with the integers $\{1,2,\ldots, 2n\}$.  Such a labeling induces a canonical labeling on $T(A)$ as is illustrated in Figure \ref{fig:labeling}.  Note that the parities of the vertex labels and of the edge labels are interchanged by $T$.  In what follows, let $\join{i}{j}$ denote the line between two points labeled $i$ and $j$ and let $\meet{k}{l}$ denote the point of intersection of two lines labeled $k$ and $l$.
	
	\begin{figure} 
%\psset{unit=1.5cm}
\begin{pspicture}(6,4)
	\rput(0,-0.5){
  % Original Pentagon
  \pspolygon[showpoints=true,linewidth=2pt](1,1)(4,1)(5,3)(3,4)(1,3)
	\uput[l](1,2){$1$}
  \uput[225](1,1){$2$}
	\uput[d](2.5,1){$3$}
  \uput[270](4,1){$4$}
	\uput[dr](4.5,2){$5$}
  \uput[0](5,3){$6$}
	\uput[ur](4,3.5){$7$}
  \uput[90](3,4){$8$}
	\uput[ul](2,3.5){$9$}
  \uput[l](1,3){$10$}
  % New Pentagon
  \pspolygon[linestyle=dashed](1,1)(5,3)(1,3)(4,1)(3,4)
  \pspolygon[showpoints=true,linewidth=2pt](1.92,2.38)(2.71,1.86)(3.57,2.28)(3.33,3)(2.33,3)
  \uput[l](1.92,2.38){$1$}
	\uput[dl](2.31,2.12){$2$}
	\uput[d](2.71,1.86){$3$}
	\uput[dr](3.14,2.07){$4$}
  \uput[330](3.57,2.28){$5$}
	\uput[20](3.45,2.64){$6$}
  \uput[ur](3.33,3){$7$}
	\uput[u](2.83,3){$8$}
  \uput[ul](2.33,3){$9$}
	\uput[160](2.12,2.69){$10$}
  
  }
\end{pspicture}
%\psset{unit=1cm}
\caption{Possible labelings of two polygons related by the pentagram map}
\label{fig:labeling}
\end{figure}

\begin{definition}
	The \emph{$y$-parameters} of a polygon $A$ are real numbers $y_1(A),\ldots, y_{2n}(A)$ defined by
	\begin{align*}
		y_i(A) = \begin{cases}
		-\left(\chi(\join{i}{(i-4)}, i-1, i+1, \join{i}{(i+4)})\right)^{-1} & \textrm{ if $i$ is a vertex of $A$} \\
		-\chi(\meet{i}{(i-4)}, i-1, i+1, \meet{i}{(i+4)}) & \textrm{ if $i$ is a side of $A$} \\
		\end{cases}
	\end{align*}
	This definition is illustrated in Figure \ref{fig:defy}.
\end{definition}

\begin{figure}
\psset{unit=.8cm}
\begin{pspicture}(12,6)
\rput(0,.5){
\psline(1,1)(3,1)(5,2)(5,3)(4,4)
\uput[d](1,1){$1$}
\uput[d](2,1){$2$}
\uput[d](3,1){$3$}
\uput[dr](4,1.5){$4$}
\uput[dr](5,2){$5$}
\uput[r](5,2.5){$6$}
\uput[r](5,3){$7$}
\uput[ur](4.5,3.5){$8$}
\uput[ur](4,4){$9$}
\psline[linewidth=2pt](1,1)(5,2)
\psline[linewidth=2pt](3,1)(5,2)
\psline[linewidth=2pt](5,3)(5,2)
\psline[linewidth=2pt](4,4)(5,2)
}
\rput(7,0){
\psline(1,1)(3,1)(5,2)(5,3)(4,4)(2,5)
\uput[d](1,1){$1$}
\uput[d](2,1){$2$}
\uput[d](3,1){$3$}
\uput[345](4,1.5){$4$}
\uput[r](5,2){$5$}
\uput[r](5,2.5){$6$}
\uput[r](5,3){$7$}
\uput[dl](4.5,3.5){$8$}
\uput[u](4,4){$9$}
\uput[u](3,4.5){$10$}
\uput[u](2,5){$11$}
\psline(3,1)(5,1)(5,3.5)(4,4)
\psdots[dotsize=2pt 3](5,1)(5,2)(5,3)(5,3.5)
}
\end{pspicture}
\psset{unit=1cm}
\caption{The cross ratios corresponding to the two types of $y$-parameters.  On the left, $-(y_5)^{-1}$ is the cross ratio of the lines $\join{5}{1}$, $4$, $6$, and $\join{5}{9}$.  On the right, $-y_6$ is the cross ratio of the points $\meet{6}{2}$, $5$, $7$, and $\meet{6}{10}$ .} 
\label{fig:defy}
\end{figure}
	
	\begin{proposition}
		Let $A$ be an $n$-gon with $y$-parameters $y_i = y_i(A)$.  Let $B= T(A)$ and denote its $y$-parameters $y_i' = y_i(B)$.  Then
		\begin{equation} \label{eq:T}
			y_i' = \begin{cases}
			y_i^{-1} & \textrm{if $i$ is a side of $B$} \\
			y_i\frac{(1+y_{i-1})(1+y_{i+1})}{(1+y_{i-3}^{-1})(1+y_{i+3}^{-1})} & \textrm{if $i$ is a vertex of $B$}
			\end{cases}
		\end{equation}
		In this equation, all indices are considered modulo $2n$.
	\end{proposition}
	
	The formula \eqref{eq:T} bares a distinct resemblance to \eqref{eq:YSystemCases}, so it is plausible that it can also be described by $Y$-pattern mutations for an appropriate quiver.  The desired quiver $Q_n$ has vertex set $\{1,2,\ldots,2n\}$ and arrows $j \to (j\pm1)$ and $j \leftarrow (j \pm 3)$ for each odd $j$, with vertices considered modulo $2n$.  The quiver $Q_8$ is in Figure \ref{fig:GlickQuiver}.  Similar to before, define compound mutations
	\begin{align*}
		\mu_{\textrm{odd}} &= \mu_{2n-1} \circ \cdots \circ \mu_3 \circ \mu_1 \\
		\mu_{\textrm{even}} &= \mu_{2n} \circ \cdots \circ \mu_4 \circ \mu_2 
	\end{align*}
	
	\begin{figure}
\begin{pspicture}(-5,-5)(5,5)
\SpecialCoor
\cnode(2;22.5){.25}{v1}
\rput(v1){1}
\cnode(4;45){.25}{v2}
\rput(v2){2}
\cnode(2;67.5){.25}{v3}
\rput(v3){3}
\cnode(4;90){.25}{v4}
\rput(v4){4}
\cnode(2;112.5){.25}{v5}
\rput(v5){5}
\cnode(4;135){.25}{v6}
\rput(v6){6}
\cnode(2;157.5){.25}{v7}
\rput(v7){7}
\cnode(4;180){.25}{v8}
\rput(v8){8}
\cnode(2;202.5){.25}{v9}
\rput(v9){9}
\cnode(4;225){.25}{v10}
\rput(v10){10}
\cnode(2;247.5){.25}{v11}
\rput(v11){11}
\cnode(4;270){.25}{v12}
\rput(v12){12}
\cnode(2;292.5){.25}{v13}
\rput(v13){13}
\cnode(4;315){.25}{v14}
\rput(v14){14}
\cnode(2;337.5){.25}{v15}
\rput(v15){15}
\cnode(4;0){.25}{v16}
\rput(v16){16}

\psset{arrowsize=5pt}
\psset{arrowinset=0}
\ncline{->}{v1}{v16}   \ncline{->}{v1}{v2}
\ncline{->}{v3}{v2}    \ncline{->}{v3}{v4} 
\ncline{->}{v5}{v4}    \ncline{->}{v5}{v6}
\ncline{->}{v7}{v6}    \ncline{->}{v7}{v8}
\ncline{->}{v9}{v8}    \ncline{->}{v9}{v10}
\ncline{->}{v11}{v10}  \ncline{->}{v11}{v12}
\ncline{->}{v13}{v12}  \ncline{->}{v13}{v14}
\ncline{->}{v15}{v14}  \ncline{->}{v15}{v16}
\ncline{->}{v16}{v13} \ncline{->}{v16}{v3}
\ncline{->}{v2}{v15} \ncline{->}{v2}{v5}
\ncline{->}{v4}{v1} \ncline{->}{v4}{v7}
\ncline{->}{v6}{v3} \ncline{->}{v6}{v9}
\ncline{->}{v8}{v5} \ncline{->}{v8}{v11}
\ncline{->}{v10}{v7} \ncline{->}{v10}{v13}
\ncline{->}{v12}{v9} \ncline{->}{v12}{v15}
\ncline{->}{v14}{v11} \ncline{->}{v14}{v1}

\end{pspicture}
\caption{The quiver related to the action of the pentagram map on octagons} \label{fig:GlickQuiver}
\end{figure}

\begin{theorem} \label{thm:pentagram}
	Let $A$ be an $n$-gon with $y$-parameters $y_1,\ldots, y_{2n}$ and let $k$ be a positive integer.  Beginning from the $Y$-seed $((y_1,\ldots, y_{2n}), Q_n)$, apply $k$ compound mutations alternating between $\mu_{\textrm{even}}$ and $\mu_{\textrm{odd}}$.  The result will have the form $((y_1',\ldots, y_{2n}'), (-1)^kQ_n)$ where 
	\begin{displaymath}
	y_i' = y_i(T^k(A)). 
	\end{displaymath}
\end{theorem}

\begin{remark}
	It is natural to ask what values the $n$-tuple $(y_1(A),\ldots, y_{2n}(A))$ can take, and also, to what extent this data suffices to reconstruct $A$.  These questions are easier to answer once extending to a larger family of objects called twisted polygons.  In this setting, the $y_i$ satisfy a single relation $y_1\cdots y_{2n}=1$.  Moreover, the $y_i$ determine a twisted polygon uniquely up to projective equivalence and a one-parameter rescaling operation due to Schwartz \cite{S08}.  The $Y$-pattern dynamics described in Theorem \ref{thm:pentagram}, then, characterize the dynamics of the pentagram map on the space of twisted polygons modulo these equivalences.
\end{remark}
	
	We now give a somewhat informal presentation of integrability of the pentagram map from the cluster algebra perspective.  The first ingredient is a compatible Poisson structure, which can be obtained directly from the $Y$-pattern analogue of Theorem \ref{thm:Poisson}.
	
	\begin{proposition}[\cite{OST10, GSTV12}]
		Define a Poisson bracket by $\{y_i,y_j\} = b_{ij}y_iy_j$ where $B$ is the exchange matrix associated to the pentagram quiver $Q_n$.  Then this bracket is preserved by the pentagram map, i.e.
		\begin{displaymath}
			\{f,g\} = \{f\circ T, g \circ T\}
		\end{displaymath}
		for all functions $f$ and $g$ of the $y$-parameters.
	\end{proposition}
	
	Some more work is needed to get at the conserved quantities.  We follow the approach of \cite{GK13}.  The quiver $Q_n$ can be lifted to an infinite doubly periodic quiver in the plane as in Figure \ref{fig:quiverLift}.  This lift represents an embedding of $Q_n$ on a torus.  Let $G = (V,E)$ be the dual graph on the torus with $V = \{1,2,\ldots, 2n\}$ where $i \in V$ corresponds to the face of the quiver with vertices $i \pm 2, i \pm 1$.  Coincidentally, $G$ is identical to $Q_n$ except without orientations, so $\overline{ij} \in E$ if and only if $i-j \equiv \pm 1, \pm 3 \pmod{2n}$.  Figure \ref{fig:torus} shows the graph $G$ on the torus in the case $n=4$, with the vertices $1,2,\ldots, 8$ appearing in order from left to right.
	
	\begin{figure}
	\begin{pspicture}(0,.5)(10,3)
	\rput(1,1){\rnode{v1}{$1$}}
	\rput(2,1){\rnode{v2}{$2$}}
	\rput(3,1){\rnode{v3}{$3$}}
	\rput(4,1){\rnode{v4}{$4$}}
	\rput(5,1){\rnode{v5}{$5$}}
	\rput(6,1){\rnode{v6}{$6$}}
	\rput(7,1){\rnode{v7}{$7$}}
	\rput(8,1){\rnode{v8}{$8$}}
	\rput(9,1){\rnode{v9}{$1$}}
	\rput(1,2){\rnode{w1}{$4$}}
	\rput(2,2){\rnode{w2}{$5$}}
	\rput(3,2){\rnode{w3}{$6$}}
	\rput(4,2){\rnode{w4}{$7$}}
	\rput(5,2){\rnode{w5}{$8$}}
	\rput(6,2){\rnode{w6}{$1$}}
	\rput(7,2){\rnode{w7}{$2$}}
	\rput(8,2){\rnode{w8}{$3$}}
	\rput(9,2){\rnode{w9}{$4$}}
	\rput(1,3){\rnode{x1}{$7$}}
	\rput(2,3){\rnode{x2}{$8$}}
	\rput(3,3){\rnode{x3}{$1$}}
	\rput(4,3){\rnode{x4}{$2$}}
	\rput(5,3){\rnode{x5}{$3$}}
	\rput(6,3){\rnode{x6}{$4$}}
	\rput(7,3){\rnode{x7}{$5$}}
	\rput(8,3){\rnode{x8}{$6$}}
	\rput(9,3){\rnode{x9}{$7$}}
	\ncline{->}{v1}{v2}
	\ncline{->}{v3}{v2}
	\ncline{->}{v3}{v4}
	\ncline{->}{v5}{v4}
	\ncline{->}{v5}{v6}
	\ncline{->}{v7}{v6}
	\ncline{->}{v7}{v8}
	\ncline{->}{v9}{v8}
	\ncline{<-}{w1}{w2}
	\ncline{<-}{w3}{w2}
	\ncline{<-}{w3}{w4}
	\ncline{<-}{w5}{w4}
	\ncline{<-}{w5}{w6}
	\ncline{<-}{w7}{w6}
	\ncline{<-}{w7}{w8}
	\ncline{<-}{w9}{w8}
	\ncline{->}{x1}{x2}
	\ncline{->}{x3}{x2}
	\ncline{->}{x3}{x4}
	\ncline{->}{x5}{x4}
	\ncline{->}{x5}{x6}
	\ncline{->}{x7}{x6}
	\ncline{->}{x7}{x8}
	\ncline{->}{x9}{x8}
	\ncline{<-}{w1}{w2}
	\ncline{->}{w1}{v1}
	\ncline{<-}{w2}{v2}
	\ncline{->}{w3}{v3}
	\ncline{<-}{w4}{v4}
	\ncline{->}{w5}{v5}
	\ncline{<-}{w6}{v6}
	\ncline{->}{w7}{v7}
	\ncline{<-}{w8}{v8}
	\ncline{->}{w9}{v9}
	\ncline{->}{w1}{x1}
	\ncline{<-}{w2}{x2}
	\ncline{->}{w3}{x3}
	\ncline{<-}{w4}{x4}
	\ncline{->}{w5}{x5}
	\ncline{<-}{w6}{x6}
	\ncline{->}{w7}{x7}
	\ncline{<-}{w8}{x8}
	\ncline{->}{w9}{x9}
	
	
\end{pspicture}
	\caption{A lift of $Q_4$ to a doubly periodic quiver in the plane.}
	\label{fig:quiverLift}
	\end{figure}
	
	\begin{figure}
	\begin{pspicture}(-1,0)(10,2)
\cnode(1,1){.08}{v1}
\cnode*(2,1){.08}{v2}
\cnode(3,1){.08}{v3}
\cnode*(4,1){.08}{v4}
\cnode(5,1){.08}{v5}
\cnode*(6,1){.08}{v6}
\cnode(7,1){.08}{v7}
\cnode*(8,1){.08}{v8}
\pnode(.5,1){w1}
\pnode(8.5,1){e8}
\pnode(0,1.5){n0}
\pnode(1,1.5){n1}
\pnode(2,1.5){n2}
\pnode(3,1.5){n3}
\pnode(4,1.5){n4}
\pnode(5,1.5){n5}
\pnode(6,1.5){n6}
\pnode(7,1.5){n7}
\pnode(8,1.16){n8}
\pnode(9,.83){n9}
\pnode(0,1.16){s0}
\pnode(1,.83){s1}
\pnode(2,.5){s2}
\pnode(3,.5){s3}
\pnode(4,.5){s4}
\pnode(5,.5){s5}
\pnode(6,.5){s6}
\pnode(7,.5){s7}
\pnode(8,.5){s8}
\pnode(9,.5){s9}
\pspolygon[linestyle=dashed](-1,1.5)(7,1.5)(10,.5)(2,.5)
\ncline{w1}{v1}
\ncline{v1}{v2}
\ncline{v2}{v3}
\ncline{v3}{v4}
\ncline{v4}{v5}
\ncline{v5}{v6}
\ncline{v6}{v7}
\ncline{v7}{v8}
\ncline{v8}{e8}
\ncline{v1}{n1}
\ncline{v2}{n2}
\ncline{v3}{n3}
\ncline{v4}{n4}
\ncline{v5}{n5}
\ncline{v6}{n6}
\ncline{v7}{n7}
\ncline{v8}{n8}
\ncline{v1}{s1}
\ncline{v2}{s2}
\ncline{v3}{s3}
\ncline{v4}{s4}
\ncline{v5}{s5}
\ncline{v6}{s6}
\ncline{v7}{s7}
\ncline{v8}{s8}
\ncline{s0}{n0}
\ncline{s9}{n9}
\end{pspicture}
	\caption{A bipartite graph on the torus dual to the quiver $Q_4$.}
	\label{fig:torus}
	\end{figure}
	
	Assign edge weights to $G$ as follows: each ``horizontal'' edge $\overline{i(i+1)}$ gets weight $1$ and each ``vertical edge'' $\overline{(i-2)(i+1)}$ gets weight $(-1)^ix_i$.  The $x_i$ represent the \emph{corner invariants} \cite{S08} of a polygon, certain coordinates related to the $y$-parameters by
	\begin{displaymath}
		y_i = \begin{cases}
		-(x_ix_{i+1})^{-1} & \textrm{if $i$ is a vertex of $A$} \\
		-x_ix_{i+1} & \textrm{if $i$ is a side of $A$} \\
		\end{cases}
	\end{displaymath}
	
	
	A \emph{perfect matching} of $G$ is a collection $M$ of its edges such that each $i \in V$ is an endpoint of exactly one $e \in M$.  The \emph{weight} of $M$, denoted $wt(M)$, is the product of the weights of its edges.  For instance, 
	\begin{displaymath}
		M = \{\overline{12}, \overline{36}, \overline{47}, \overline{58}\}
	\end{displaymath}
	is a perfect matching of $G_4$ and its weight is $(1)(-x_5)(x_6)(-x_7) = x_5x_6x_7$.  Finally, there is a notion of homology class of a perfect matching, which we do not describe here.
	
	\begin{proposition}
		For any homology class, the sum
		\begin{displaymath}
			\sum_M wt(M)
		\end{displaymath}
		over perfect matchings $M$ of $G_n$ of that class is a conserved quantity of the pentagram map acting on $n$-gons.	\end{proposition}
	
	In the case of the graph $G_4$ pictured in Figure \ref{fig:torus}, there are $22$ perfect matchings which represent $8$ different homology classes.  Table \ref{table:integrals} gives the $8$ corresponding conserved quantities for the pentagram map.  The first $2$ are trivial but the remaining $6$ are (up to a sign) the original conserved quantities discovered by Schwartz \cite{S08} by different means.
	
	\begin{table}[ht]
	\begin{tabular}{l|l|l}
	Name & Representative matching & Formula \\
	\hline
	n.a.  & $\{\overline{12}, \overline{34}, \overline{56}, \overline{78}\}$ & $1$\\
	n.a.  & $\{\overline{23}, \overline{45}, \overline{67}, \overline{18}\}$ & $1$\\
	$O_1$ & $\{\overline{27}, \overline{34}, \overline{56}, \overline{18}\}$ 
	& $-x_1-x_3-x_5-x_7+x_1x_2x_3+x_3x_4x_5+x_5x_6x_7+x_1x_7x_8$ \\
	$E_1$ & $\{\overline{38}, \overline{12}, \overline{45}, \overline{67}\}$
	& $x_2+x_4+x_6+x_8-x_2x_3x_4-x_4x_5x_6-x_6x_7x_8-x_1x_2x_8$ \\
	$O_2$ & $\{\overline{27}, \overline{36}, \overline{45}, \overline{18}\}$
	& $x_1x_5 + x_3x_7$\\
	$E_2$ & $\{\overline{38}, \overline{47}, \overline{12}, \overline{56}\}$
	& $x_2x_6 + x_4x_8$ \\
	$O_4$ & $\{\overline{14}, \overline{36}, \overline{58}, \overline{27}\}$
	& $x_1x_3x_5x_7$ \\
	$E_4$ & $\{\overline{16}, \overline{38}, \overline{25}, \overline{47}\}$
	& $x_2x_4x_6x_8$ \\
	\end{tabular}
	\caption{Conserved quantities for the pentagram map acting on twisted quadrilaterals.  The matching $M$ is such that $wt(M)$ is the first term appearing in the formula.}
	\label{table:integrals}
	\end{table}
	
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{thebibliography}{99}
  \bibitem{ABCGPT14} N. Arkani-Hameda, J. Bourjailyb, F. Cachazoc, A. Goncharovd, A. Postnikove, and J. Trnka, Scattering amplitudes and the positive Grassmannian, arXiv preprint: 1212.5605v2.
  \bibitem{BFZ05} A. Berenstein, S. Fomin, and A. Zelevinsky, Cluster algebras III: upper bounds and double Bruhat cells, \textsl{J. London Math. Soc.} \textbf{14} (1976), 183-187.
  \bibitem{BR15} A. Berenstein and D. Rupel, Quantum cluster characters of Hall algebras, \textsl{Sel. Math. New Ser.} \textbf{21} (2015), no. 4, 1121--1176.
  \bibitem{CC06} P. Caldero and F. Chapoton, Cluster algebras as Hall algebras of quiver representations, \textsl{Comment. Math. Helv.} \textbf{81} (2006), no. 3, 595--616.
  \bibitem{CK06} P. Caldero and B. Keller, From triangulated categories to cluster algebras II, \textsl{Ann. Sci. \'Ecole Norm. Sup.} (4) \textbf{39} (2006), no. 6, 983--1009.
  \bibitem{BMRRT06} A. Buan, R. Marsh, I. Reiten, M. Reineke, and G. Todorov, Tilting theory and cluster combinatorics, \textsl{Adv. Math.} \textbf{204} (2006), no. 2, 572-618.
	\bibitem{dFK10} P. DiFrancesco and R. Kedem, $Q$-systems, heaps, paths and cluster positivity, \textsl{Comm. Math. Phys.} \textbf{293} (2010), 727--802.
  \bibitem{EF12} R. Eager and S. Franco, Colored BPS pyramid partition functions, quivers and cluster transformations, \textsl{J. High Energy Phys.} (2012), no. 9, 038.
	\bibitem{FG06} V. Fock and A. Goncharov, Moduli spaces of local systems and higher Teichmller theory, \textsl{Publ. Math. Inst. Hautes tudes Sci.} \textbf{103} (2006), 1--211.
	\bibitem{FST08} S. Fomin, M. Shapiro, and D. Thurston, Cluster algebras and triangulated surfaces. I. Cluster complexes, \textsl{Acta Math.} \textbf{201} (2008), 83--146.
	\bibitem{FZ02} S. Fomin and A. Zelevinsky, Cluster algebras I: Foundations, \textsl{J. Amer. Math. Soc.} \textbf{15} (2002), 497--529.
  \bibitem{FZ03} S. Fomin and A. Zelevinsky, Cluster algebras II: Finite type classification, \textsl{Invent. Math.} \textbf{154} (2003), 63-121.
  \bibitem{FZ02b} S. Fomin and A. Zelevinsky, The Laurent phenomenon, \textsl{Adv. in Applied Math.} \textbf{28} (2002), 119-144.
	\bibitem{FZ07} S. Fomin and A. Zelevinsky, Cluster algebras IV: Coefficients, \textsl{Compos. Math.} \textbf{143} (2007), 112--164.
	\bibitem{FZ99} S. Fomin and A. Zelevinsky, Double Bruhat celles and total positivity, \textsl{J. Amer. Math. Soc.} \textbf{12} (1999), 335--380.
	\bibitem{GLS06} C. Geiss, B. Leclerc, and J. Schroer, Rigid modules over preprojective algebras, \textsl{Invent. Math} \textbf{165} (2006), 589--632.
  \bibitem{GLS13} C. Geiss, B. Leclerc, and J. Schr\"oer, Cluster structures on quantum coordinate rings, \textsl{Selecta Math. (N.S.)} \textbf{19} (2013), no. 2, 337--397.
  \bibitem{GLS} C. Geiss, B. Leclerc, and J. Schr\"oer,
	\bibitem{GSV03} M. Gekhtman, M. Shapiro, and A. Vainshtein,  Cluster algebras and Poisson geometry, \textsl{Mosc. Math. J.} \textbf{3} (2003), 899--934.
  \bibitem{GSV10} M. Gekhtman, M. Shapiro, and A. Vainshtein,  Cluster algebras and Poisson geometry, Mathematical Surveys and Monographs, \textbf{167}. American Mathematical Society, Providence, RI, 2010.
	\bibitem{GSTV12} M. Gekhtman, M. Shapiro, S. Tabachnikov, and A. Vainshtein, Higher pentagram maps, weighted directed networks, and cluster dynamics, \textsl{Electron. Res. Announc. Math. Sci.} \textbf{19} (2012), 1--17.
	\bibitem{G11} M. Glick, The pentagram map and Y-patterns, \textsl{Adv. Math.} \textbf{227} (2011), 1019--1045.
	\bibitem{GK13} A. Goncharov and R. Kenyon, Dimers and cluster integrable systems, \textsl{Ann. Sci. \'Ec. Norm. Sup\'er.} \textbf{46} (2013), 747--813.
  \bibitem{K10} B. Keller, Categorification of acyclic cluster algebras: an introduction, arXiv preprint: 0801.3103v2.
  \bibitem{K12} Y. Kimura, Quantum unipotent subgroup and dual canonical basis, \textsl{Kyoto J. Math.} \textbf{52} (2012), no. 2, 277--331.
  \bibitem{KQ14} Y. Kimura and F. Qin, Graded quiver varieties, quantum cluster algebras and dual canonical basis, \textsl{Adv. Math.} \textbf{262} (2014), 261--312.
  \bibitem{KZ02} M. Kogan and A. Zelevinsky, On symplectic leaves and integrable systems in standard complex semisimple Poisson-Lie groups, arXiv preprint: 0203069v1.
  \bibitem{LLZ14} K. Lee, L. Li, and A. Zelevinsky, Greedy elements in rank 2 cluster algebras, \textsl{Sel. Math. New Ser.} \textbf{20} (2014), 57--82.
  \bibitem{LS13} K. Lee and R. Schiffler, A Combinatorial Formula for Rank 2 Cluster Variables, \textsl{J. Alg. Comb} \textbf{37} (2013), 67--85.
  \bibitem{LS15} K. Lee and R. Schiffler, Positivity for cluster algebras, \textsl{Ann. Math.} \textbf{182} (2015), 73--125.
	\bibitem{MP07} G. Musiker and J. Propp, Combinatorial interpretations for rank-two cluster algebras of affine type, \textsl{ Electron. J. Combin.} \textbf{14} (2007), 23 pp.
	\bibitem{MSW13} G. Muisker, R. Schiffler, and L. Williams, Bases for cluster algebras from surfaces, \textsl{Compos. Math} \textbf{149} (2013), 2891--2944.
	\bibitem{OST10} V. Ovsienko, R. Schwartz, and S. Tabachnikov, The pentagram map: a discrete integrable system, \textsl{Comm. Math. Phys.} \textbf{299} (2010), 409-446.
	\bibitem{OST13} V. Ovsienko, R. Schwartz, and S. Tabachnikov, Liouville-Arnold integrability of the pentagram map on closed polygons, \textsl{Duke Math. J.} \textbf{162} (2013), 2149--2196.
  \bibitem{Q12} F. Qin, Quantum cluster variables via Serre polynomials, \textsl{J. Reine Angew. Math.} \textbf{668} (2012), 149--190.
  \bibitem{R11} D. Rupel, On a quantum analogue of the Caldero-Chapoton formula, \textsl{Int. Math. Res. Not.} (2011), no. 14, 3207--3236.
  \bibitem{R15} D. Rupel, Quantum cluster characters for valued quivers, \textsl{Trans. Amer. Math. Soc.} \textbf{367} (2015), 7061--7102.
	\bibitem{S92} R. Schwartz, The pentagram map, \textsl{Experiment. Math.} \textbf{1} (1992), 71--81.
	\bibitem{S08} R. Schwartz, Discrete monodromy, pentagrams, and the method of condensation, \textsl{J. Fixed Point Theory Appl.} \textbf{3} (2008), 379--409.
	\bibitem{S13} F. Soloviev, Integrability of the Pentagram Map, \textsl{Duke Math. J.} \textbf{162} (2013), 2815--2853.
	\bibitem{V07} A. Volkov, On the periodicity conjecture for $Y$-systems, \textsl{Comm. Math. Phys.} \textbf{276} (2007), 509--517.
  \bibitem{W13} L. Williams, Cluster algebras: an introduction, arXiv preprint: 1212.6263v3.
\end{thebibliography}

\end{document}